Ab initio prediction of specific phospholipid complexes and membrane association of HIV-1 MPER antibodies by multi-scale simulations

Version 1

Abstract

Summary

A potent class of HIV-1 broadly neutralizing antibodies (bnAbs) targets the envelope glycoprotein’s membrane proximal exposed region (MPER) through a proposed mechanism where hypervariable loops embed into lipid bilayers and engage headgroup moieties alongside the epitope. We address the feasibility and determinant molecular features of this mechanism using integrative modeling. All-atom simulations of 4E10, PGZL1, 10E8 and LN01 docked onto HIV-like membranes consistently form phospholipid complexes at key complementarity-determining region loop sites, solidifying that stable and specific lipid interactions anchor bnAbs to membrane surfaces. Ancillary protein-lipid contacts reveal surprising contributions from antibody framework regions. Coarse-grained simulations effectively capture antibodies embedding into membranes. Simulations estimating protein-membrane interaction strength for PGZL1 variants along an inferred maturation pathway show bilayer affinity is evolved and correlates with neutralization potency. The modeling platform developed here uncovers insights into lipid participation in antibodies’ recognition of membrane proteins and highlights antibody features to prioritize in vaccine design.

Introduction

Antibodies can target epitopes on integral membrane proteins very near to the lipid bilayer surface, even those partially embedded within the headgroup region. In the process of engaging the antigen, antibodies complementarity-determining regions (CDR) may need to navigate interactions with or concealment by lipid molecules. Then, evolving tolerance or even affinity to lipid bilayers could be beneficial in recognition of membrane-proximal epitopes, for example gaining avidity and specificity via cooperative interactions simultaneously with the epitope and membrane. Conversely, propensity to bind lipids or cell membranes poses a significant auto-immunity risk. B-cells producing antibodies targeting host membranes are downregulated in healthy organisms1–4. Nonetheless, in cases of chronic inflammation and infection as in HIV, rare antibodies with lipid affinity can emerge5–8. The maturation pathways of these rare events remain unclear, but a careful balance of polyreactivity must be struck to avoid or overcome autoreactivity9. A better understanding of how antibodies develop membrane affinity and target membrane-proximal epitopes would be impactful for antibody therapeutics, auto-immunity, and vaccine development10–12.

We sought to address this phenomenon for broadly neutralizing antibodies (bnAbs) 4E10, PGZL1, 10E8, and LN01 of unique lineages which all target the semi-concealed membrane-proximal epitope region (MPER)13–16 of the HIV-1 envelope glycoprotein (Env). Interestingly, most of these bnAbs show affinity for lipid components and freely associate with lipid bilayers5–8 or cultured cells, even in the absence of antigen17–23. This membrane interaction behavior is attributed to shared CDR loop features, including a long hydrophobic CDR- H37, 18, 21, 22, 24–26, and appears to correlate with neutralization potency. Engineered mutations to the 10E8 paratope at light-chain solvent-exposed residues that add positive charge or hydrophobic sidechains proved to increase antibody association to anionic phospholipid vesicles in vitro and also boosted neutralization potency27. Conversely, mutations reducing CDR-H3 hydrophobicity (e.g. 4E10 H100-H102 Trp-Trp motif to Ala-Ala or Asp-Asp) only marginally affect antigen affinity, but drastically reduce neutralization activity and weakened association to lipid bilayers in parallel7, 21, 28, 29.

Structural characterization of full-length Env trimer or fragments of gp41 suggest the surrounding lipid bilayer plays a role in antibody access and epitope recognition13, 30. Cryo-electron microscopy (cryo-EM) of bnAb-bound pre-fusion Env in different membrane environments indicate that bnAb CDR loops form intimate contacts with surrounding lipids while engaging MPER. Additionally, crystal structures of MPER-targeting bnAbs as antigen binding fragments (Fab) soaked with short-chain phospholipids revealed a spectrum of ordered headgroup moieties complexed within the CDR loops in the presence and absence of antigen, suggesting the antibodies encode specific lipid interactions5, 8, 20. Thus, the molecular features mediating membrane affinity for these bnAbs appear critical to their maturation and mechanism of immune protection against HIV in vivo.

These data support a 2-step bnAb neutralization mechanism proposed previously, wherein a population of bnAbs in vivo may first associate with membranes via embedding their CDRs, then laterally diffuse across the bilayer surface to subsequently engage Env at MPER6, 19, 22, 24, 25, 30–33. Here, we developed multi-scale molecular dynamics (MD) simulation approaches suited to investigate this mechanism alongside the unique maturation landscape these rare antibodies must navigate to avoid auto-immune consequences. For these MPER bnAbs, we focused on the ability to model and characterize in silico the molecular features that mediate their lipid affinity in vivo at the necessary atomic detail not afforded by structural approaches used to date. We expect these and similar approaches for predicting membrane bound conformations and lipid affinity of natural and putative antibody sequences will be powerful for understanding neutralization efficacy, epitope engagement mechanisms, and the checkpoints regulating self-sensing antibodies.

Modelling bnAbs as Fabs docked onto HIV-like lipid bilayers, we find that unbiased all-atom MD simulations accurately and reproducibly predict ab inito binding of phospholipids at specific CDR binding sites previously identified in co-crystal structures19, 20, 24, 31. Further, we demonstrate that macroscopic surface properties represented in unbiased insertion coarse grain (CG) simulations are sufficient to capture the antibodies’ propensities to associate with membranes. The globally favored membrane-inserted Fab geometries agree across our multi-scale simulation stages and experiments. Extending CG trajectories into all-atom simulations facilitated building a thorough integrative model of bnAbs docking onto membranes to target Env. That is, bnAbs initially associate with a membrane surface, progress to a low energy CDR-embedded conformational ensemble wherein stable and specific CDR-mediated phospholipid complexes form, and then dynamically re-orient for MPER binding. Overall, the atomic simulations illuminate key molecular features that encode and tune each bnAbs’ range of preferred membrane-bound geometries. These molecular details include surprisingly important and underappreciated contributions from framework regions. To approximate membrane interaction strength of antibody models in silico, we benchmarked all-atom constant velocity pulling simulations in explicit lipid bilayers on 4E10 variants, successful differentiating experimentally determined differences in membrane binding affinity21, 25, 28. Finally, we tracked the membrane association strength of PGZL1 germline-reverted variants to mimic a pseudo-maturation pathway, demonstrating significant increase in lipid interactions as the antibody matures, which correlates with experimental neutralizing e fficacy18.

We demonstrate the utility of this novel modeling framework to retrospectively study or potentially proactively engineer antibodies to better navigate the proposed 2-step bnAb neutralization mechanism and facilitate HIV vaccine design. Likewise, our results have broad implications for the in vivo selection of antibodies targeting membrane-embedded proteins. We expect the insights and approaches highlighted here to translate generally, beyond HIV, towards generalized principles for targeting membrane proteins at partially concealed juxtamembrane regions, which is highly desirable for therapeutic antibodies and vaccine design given the high protein sequence conservation often found near or within the membrane.

Atomic simulations accurately model interactions between MPER bnAbs and HIV-like lipid bilayers

To assess the stability, organization, and spectrum of phospholipid interactions inferred from previous lipid-bound crystallography and cryo-EM experiments for these bnAbs, we first performed unbiased all-atom simulations of 4E10, PGZL1, and 10E8 peripherally embedded to the surface of model lipid bilayers. Starting membrane-bound conformations for each Fab were computed using a structure-based docking method, PPM2.034, globally optimizing insertion using a per-residue solvation pseudo-energy based on residue hydrophobicity. All bnAbs were reasonably predicted with their CDR-H3 inserted into the membrane. Explicit solvent systems were built with simplified HIV-like anionic cholesterol-rich membranes (25% cholesterol, 5% 1-palmitoyl-2-oleoyl- sn-glycero-3-phosphate (POPA), 70% 1-palmitoyl-2. -oleoyl-sn-glycero-3-phosphocholine (POPC)), and four pseudo-replicate simulations of 1 microsecond (μs) each were initiated for each bnAb, with two replicates tilted by ±15 degrees to modestly vary the starting membrane-interacting pose.

In each pseudo-replicate simulation of each antibody, we observed one phospholipid to stably associate at a single site proximal to the Fab heavy-chain CDR loops (Figures 1-2). In some cases, phospholipid binding events formed very quickly, as early as the protein-restrained equilibration (15 ns) where lipids can freely diffuse (Figure S1D,E). Replicate simulations with artificially tilted Fab still robustly formed the CDR-mediated phospholipid complexes, usually after 50-250 ns of minor Fab reorientation on the membrane surface (Supp. Video 1). These phospholipid binding events were highly stable, typically persisting for hundreds of nanoseconds. Across the three bnAbs, each respective CDR-phospholipid complex was present in >70% of the total aggregated 4 μs simulation time (Figures 1C-D, 2B, S1D-F). Time-averaged densities of the bound phospholipid atoms showed clear atomic overlap with electron density located at CDR sites within lipid-soaked X-ray structures wherein lipid headgroups were modelled (Figures 1A-B, 2A, S1A-C)20, 24, 31. Our simulations uncover the CDR-lipid interactions that are the most feasible and biologically relevant in the context of a full lipid bilayer. These results add validation to which out of the many possible ions and short-chain lipid moieties ordered within those crystal structures may be inferred as putative lipid. Notably, the loop backbone conformations for these CDRs remained internally rigid during the simulations (Figure S2A,E,H). Thus, these bnAbs constitutively present these pre-formed CDR binding sites for phospholipid loading at the bilayer surface upon insertion. In a PGZL1 simulation, we observed a lipid exchange event (Figure S1E), where a long-bound POPC molecule dissociated and was promptly replaced by a POPA molecule. Thus, while bound phospholipids are kinetically stable (∼102 ns), they can freely associate and dissociate with CDR loops within the timescale of our simulations. Together, these behaviors indicate that phospholipids likely readily exchange at these CDR loop sites in vivo for membrane-associated bnAbs, with equilibrium strongly favoring the bound state. In summary, despite the different bnAbs mediating membrane interactions via distinct protein features, our simulations robustly and reproducibility capture the phospholipid binding process ab inito at experimentally determined CDR loop sites with atomic accuracy.

4E10 and PGZL1 CDR lipid phosphate binding sites and membrane interaction profiles in all- atom fluid HIV-like bilayer MD simulations

(A) Representative frame from an MD simulation of the phospholipid complex at 4E10 Fab CDR-H1 (top left). Top right, time-averaged lipid phosphate density (orange mesh) relative to antibody CDR loops embedded in the bilayer (beige) from 4 µs total time (top right). Center right, CDR-H1 loop side chain and backbone atoms of de novo predicted phosphate interaction (middle right). Bottom, putative lipid phosphate binding site observed in an X-ray structure for 4E10 (PDB: 4XC1), comparing the CDR-H1 loop interactions to MD (bottom right).

(B) PGZL1 lipid phosphate interaction at the CDR-H1 loop from MD simulation versus X-ray crystallography (PDB: 6O3J), demarked as in (A).

(C) RMSD of the interacting lipid phosphate versus the experimental CDR-phosphate position (X-ray site), classified as bound (green) and unbound (grey) by loop-phosphate contacts in 2 4E10 representative replicate trajectories. Black line, ten-frame RMSD running average; standard deviation, grey shading.

(D) RMSD of lipid phosphate binding to PGZL1 CDR-H1 in MD simulations versus X-ray site.

(E) Per-residue interaction profiles for Fab simulations of 4E10 detailing the time spent for each residue in in phosphate layer (orange), glycerol layer (red), or hydrocarbon layer (blue) across aggregate 4 µs from 4 simulations. CDR loops are mapped in solid color blocks below each profile, and Fab domain regions making significant contact are labeled.

(F) Per-residue interaction profiles for antibody Fab simulations for PGZL1, colored as in (E).

10E8 bivalent lipid headgroup interaction and bilayer insertion predicted in MD simulation closely matches experimental lipid binding sites

(A) Top left, representative frame from MD simulation of lipid interacting with 10E8 Fab. Top right, the MD lipid binding site includes a bivalent choline and phosphate lipid headgroup complex (represented as blue and orange mesh time-averaged positional density in simulations) within a protein surface groove composed of CDR-L1 and FR-L3 respectively. Bottom, positions of phosphates or glycerols modeled at the CDR-L1 and FR-L3 groove site within 10E8 Fab X-ray structures (PDB: 5T85, 5T6L)

(B) RMSD of lipid choline position in the MD simulations versus expected CDR-L1 lipid binding site from 10E8 X-ray structures. “Bound” state assigned relative to choline position and phosphate FR-L3 interactions observed in MD.

(C) Per-residue interaction profiles for antibody Fab simulations for 10E8, colored as in Figure 1E.

4E10 and PGZL1 are structurally homologous (85% sequence identity, sharing IGHV1-69 germline), both forming complexes at their CDR-H1 loop site predominantly coordinating via the lipid phosphate group20, 31, 35. Across 4E10 and PGZL1 simulations, the bound lipid phosphates were generally rigid within the CDR site (<0.75 Å RMSF), and maintained sub-atomic mean RMSDs relative to the reference position in X-ray structures: 0.6 and 0.7 Å, respectively (Figure 1C-D, S2B,F). Likewise, the interactions coordinating the lipid phosphate oxygens at CDR-H1 were similar with atomic accuracy between MD simulations and the X-ray structures, with hydrogen bonds donated from Phe29 and Ser30 backbone amides and from Ser29/Thr29 and Ser30 sidechain hydroxyls20, 31 (Figure 1A-B, S1A-B). Although the headgroup phosphate displayed some positional dynamics and variation, this extensive polar network was mostly maintained over time – modestly fluctuating to accommodate the lipid.

These results of reliably recovering experimentally determined CDR-phospholipid complexes establish confidence that other antibody features in the simulation, namely the membrane-bound Fab conformation and ancillary protein-lipid interactions, are biologically relevant. Thus, we analyzed the additional molecular and sequence level features driving the lipid bilayer interactions. Time-averaged per-residue lipid interaction profiles separated by membrane layer (polar headgroup, glycerol, and hydrocarbon) show that 4E10 and PGZL1 heavy- chain loops were deeply immersed into the bilayer, whereas light-chains had sparse and only surface-level headgroup contacts (CDR-L1, -L2) (Figure 1E-F). 4E10 CDR-H loops buried deeper into hydrocarbon chain layer of membrane and formed more extensive contacts with the lipid aliphatic tails compared to PGZL1 CDR- H loops. These patterns may in part explain 4E10’s greater poly-reactivity. One striking and novel observation consistent between 4E10 and PGZL1 simulations was that heavy-chain framework region 3 (FR-H3) was extensively embedded, showing that surface features, which are predominantly germline encoded, contribute significantly to membrane binding and Fab orientation. Aligning interaction profiles with corresponding primary sequences for light-chain loops and FR-H3 reveals that charged residues mediate their shallower surface interactions (Figure S3A-B), whereas the more deeply inserted CDR-H loops bare neighboring stretches of hydrophobic and lipophilic polar residues. These analyses supply comprehensive atomic-detail maps of bnAb- membrane interfaces. Delineating the molecular features defining these protein-lipid surfaces adds key context to complement the battery of previous experiments characterizing the insertion, accessibility, and relative orientation of the MPER epitope with and without bnAbs – additionally highlighting contributions from framework regions to membrane interaction previously underestimated.

We similarly investigated bnAb 10E8, which differs in its genetic origins and expected light-chain- mediated membrane binding mode. In all 4 replicate simulations, we observed a POPC complexed at a groove between CDR-L1 and FR-L3 which had had significant non-protein electron density in previous 10E8 X-ray structures with and without lipids (Figures 2A-B, S1C,F), modelled as headgroup phosphoglycerol anions, glycerol, or free dihydrogen phosphate. In our simulations, the POPC phosphate was slightly offset (2.7 Å) from the proposed CDR-L1 crystallographic site, instead complexing mainly with FR-L3 stabilized by a hydrogen bonding network from Ser83 hydroxyl, Gly84 backbone amide, and Asn85 C-alpha proton (Figures 2A, S1C). In parallel, a POPC choline moiety simultaneously occupied the CDR-L1 site, paring the choline cation with four carbonyls from the CDR-L1 backbone due to its short helical conformation (Figure 2B). Thus, our simulation finds stable bivalent complexes at this CDR-L1 FR-L3 groove with the lipid phosphocholine (PC) zwitterion with >70% occupancy overall across replicate simulations (CDR-L1–choline site RMSF = 0.5 Å; FR-L3–phosphate site RMSF = 0.5 Å). These 10E8 loops were also conformationally rigid (RMSF < 1.0 Å, Figure S2H), presenting a pre-formed binding site. The bilayer in our simulations provides a more realistic and contextualized environment to model these interactions compared to lipid-soaked X-ray crystallography structures, where lipids have unlimited degrees of freedom to access loops. This CDR-L1 site highlights the importance of phospholipid binding features at CDR-L1 in the 10E8 binding and neutralization mechanism, given that the double mutant of R29 and Y32 to alanine exhibited vastly decreased neutralization (>500 fold) across HIV strains24.

Additionally, we sought to assess the biological relevance and in silico predictability of these observed CDR-lipid polar interactions using an orthogonal bioinformatics approach. We mined protein-ligand interactions to assessed whether phosphate complexes of similar geometries have been observed in nature, positing that the occurrence of analogous phosphate interactions outside the context of MPER bnAbs, i.e. within critical structural or functional regions of other protein families, can differentiate common proteome-wide functional motifs from simulation artifacts. Structural searches querying the lipid-binding loop backbone conformation of each Fab (as previously described14, 20, 24) identified between 105 to 2·106 geometrically similar sub-segments within natural proteins (<2 Å RMSD)37, highlighting their prevalence in the protein universe. These hits were subsequently mined for nearby phosphate/phosphoryl and sulfate/sulfo ligands (Figure S4). Only 4 cases of phosphate-type ligand binding motifs were identified when searching structures similar to 4E10’s and PGZL1’s CDR-H1. As such, this CDR-H1 site is a realistic but rare protein-phospho-ligand structural motif (Figure S4A,B; Table S1). For 10E8’s CDR-L1 site, only 1 phosphate-type ligand was observed: a solvent-exposed surface crystallographic ion. By contrast, 10E8’s FR-L3 beta-turn site was a hot spot for phosphate-type ligands, with 23 natural protein cases (Figure S4C, Table S1). Thus, while both CDR-L1 and FR-L3 10E8 loops sites can bind phosphate ligands, the much greater natural prevalence of phosphates ligated at FR-L3 supports that the 10E8 bivalent POPC complex in our simulations is a biologically realistic site. Barring experimental validation, existence of this FR- L3 site establishes precedent for how important framework regions can be for mediating protein-membrane interactions, including hosting stable and specific phospholipid complexes.

Next, we examined the geometries of the membrane-bound conformational ensembles sampled in 10E8, 4E10, and PGZL1 simulations, given the emphasis on tilt of the antigen-antibody complex relative to the bilayer in proposed mechanisms of neutralization30 and the potential energetic cost, if any, for reorienting to facilitate MPER access and complex with Env. Fab conformations were described by the immersion depths of each CDR loop and 2 angles relating the global domain orientation to the membrane. Structural clustering distinguished micro- and macroscopic sub-states sampled globally and individually (Figure S5A-C), with 6 or 7 relatively well- distributed micro-states for each bnAb (5 – 30 % population fractions) describing the spectrum of surface-bound conformations (Figure S4D,F,G). Interestingly, while all 4E10 micro-states maintain 86-98% occupancy of the CDR-H1 phospholipid complex, 10E8 and PGZL1 microstates showed more variance in propensity to form CDR- phospholipid complexes (Figure S5E,G,I). Namely, two of 10E8’s micro-states had low CDR-L1 PC headgroup occupancy (23% and 33%), while two other micro-states were bound in 83% and 78% of frames, respectively. Thus, formation and maintenance of key phospholipid complexes can depend on the CDR and FR binding site loop orientation and accessibility at the bilayer headgroup sector, highlighting the importance each bnAbs’ characteristic surface-bound conformational ensemble upon association and the protein sequence features mediating these geometries.

We next studied bnAb LN01, previously crystallized with a minimal MPER fragment as well as an extended TM-MPER fragment, to interrogate differences in the antibody surface-bound conformational ensemble in the context of a transmembrane antigen (Figure 3A,C)19. These MPER-bound LN01 structures also included ordered detergent and phospholipids, and suggest two putative relevant CDR-phospholipid binding sites. Site one occurs at an CDR-H3-L1-L2 inter-chain groove, coordinating a PC headgroup via a CDR-L1 Lys31 salt bridge to the lipid phosphate and engages the choline with CDR-L1 Tyr32 and CDRH3Tyr100g in a cation-pi cage motif – optionally including the MPER epitope helix Trp680 and Tyr681 aromatics (Figures 3E,G, S1G). The more solvent-exposed second site between CDR-H1 and -H2 binds phosphatidylserine (PS) or PC headgroups. In replicate half-μs simulations for both LN01 alone and the LN01-MPER-TM complex in HIV-like model membranes, several POPC binding events were observed at the CDR-H3-L1-L2 groove. The headgroup polar atom positioning in LN01’s sidechain interaction network was reproducibly recovered with atomic accuracy versus ordered dodecyl-PC to X-ray structures (Figure 3E-H). However, compared with PGZL1, 4E10, and 10E8 simulations, the kinetic stability and occupancy for phospholipids in LN01 at the CDR-H3-L1-L2 groove site were much lower ∼30% and ∼40% phospholipid-bound overall for LN01-MPER-TM and apo, respectively (Figures 3I-J, S1H-I). The second crystallographic CDR-H1-H2 PS/PC site was not recovered in any replicate simulation.

Atomistic simulations of apo and antigen-bound LN01 characterizing paratope-phospholipid complexes with and without epitope and shifted membrane-bound conformation

(A) Representative frame from MD simulation LN01 Fab bound to MPER-TM showing the stable ternary paratope-epitope-membrane complex; bound phospholipids shown.

(B) Frequency of Fab’s characteristic surface-bound geometry by global domain rotation and approach angles in MD simulations for LN01 bound to MPER-TM, plotted by kernel density estimation as contour.

(C) Representative frame from MD simulation of phospholipids complexed with LN01 Fab alone.

(D) Frequency of geometries sampled for apo membrane-bound LN01.

(E) Phospholipid headgroup interaction formed ab initio in LN01+MPER-TM simulations. Aromatic cation-pi cage motifs coordinate choline while the phosphate is coordinating by Lys31 matching the X- ray site binding pose.

(F) The additional distal “Loading” phospholipid site predicted in LN01simulations, with a similar cation-pi cage motif and hydrogen bonds interactions stabilizing the PC headgroup.

(G) Atomic interactions at the X-ray site in apo LN01 simulations.

(H) Interactions at the loading site in apo LN01 simulations.

(I) Lipid headgroup binding in representative simulation of LN01+MPER-TM (n=4 total). Top, X-ray binding site occupancy (green) and phospholipid choline RMSD in a representative trajectory versus experimental position. Bottom, loading site occupancy (cyan) and choline RMSD versus average headgroup bound position.

(J) Lipid headgroup binding for representative apo LN01 trajectory (n=4 total). Site occupancy and RMSD versus predicted binding position for X-ray site (top) and loading site (bottom).

(K) Per-residue interaction profile for MPER-TM-bound LN01.

(L) Per-residue interaction profile for apo LN01.

Two novel features stood out in LN01 simulations. First, antigen-free LN01 sampled a broader breadth of geometries at the membrane surface, characterized by two angles describing the Fab relative to the membrane (Figure 3A-D). LN01 was more deeply and extensively embedded when complexed with MPER-TM than LN01 alone. The more focused conformation landscape likely resulted from the Fab accommodating the inserted MPER-TM antigen’s lowest energy orientation. Comparing the per-residue protein-lipid interaction profiles revealed consistency in the protein molecular features mediating membrane association (Figure 3K-L), although CDR-L, CDR-H, and FR-L3 loops were inserted further upon TM antigen engagement. While the MPER-TM fragment may present the epitope differently than the full length Env trimer, our data describe how a bnAb shifts its global geometry to access and engage a transmembrane-presented MPER; yet, the apo state is predisposed based on CDR loop insertion and samples this geometry readily within its resting conformational ensemble on the membrane surface. Then, the bnAbs’ sequence and structure are primed for transmembrane epitope engagement.

The second notable feature was an additional LN01 CDR-phospholipid binding site predicted across simulation replicates at an alternative groove comprised of CDR-H3, FR-L2, and CDR-H2. We named this the “Loading site” given its proximity (∼7 Å) to the previously observed CDR-phospholipid H3-L1-L2 “X-ray” site and observations that phospholipids quickly exchange from the Loading site into the X-ray site by diffusion (Figure S1H,I). Although, the two sites were often simultaneously occupied by distinct lipids. Similar to the X- ray site, Loading site binds lipids in the apo state, but in the presence of the MPER-TM can optionally including antigen TM residue Trp680 (Figures 3F,H,I-J, S1G). LN01 Tyr100g, Tyr49, and Tyr52 (and gp41’s Trp680 in MPER-TM-bound simulations) hydrogen bond with the lipid headgroup phosphate oxygens. The POPC choline is simultaneously engaged by electrostatic contacts from Tyr49, Tyr52, Thr53, Thr100c, Ser100d, and Tyr100g side chains, forming second cation-pi cage motif structurally analogous to that used at the X-ray site (Figure 3E- H). The Loading site was occupied an average of 78% of LN01-MPER-TM simulation time versus 58% for apo LN01 replicates (Figures 3I,J S1H,I). Thus, these LN01 simulations highlight that the paratope-epitope- membrane complex robustly reforms and is stable in the lipid bilayer context. The ternary complex also may incorporate multiple phospholipid participants loaded prior to epitope engagement, which add stabilizing interactions with the MPER helix. These results strengthen the argument that antibodies can evolve structural elements to engage both a membrane protein antigen and the surrounding lipid environment – and that we can predict as well as possibly engineer these molecular features in silico.

Coarse-grain ab initio insertion simulations capture biologically relevant membrane-bound conformations

Because the limited sampling accessible in all-atom simulations, we turned to coarse-grained (CG) simulations utilizing the Martini model to explore the process of antibody insertion to lipid bilayers and more thoroughly assess the range of plausible and favorable bnAb surface-bound geometries. With elastic network restraints, CG representation was structurally stable in simulations, maintaining Fab tertiary and quaternary structure (<2 Å backbone RMSD, Figure S6A). Then, we hypothesized that this simulation model, which rigidly presents Fabs’ macroscopic surface features (hydrophobic patches, charged patches, polarity) as polar beads, would be sufficient to capture these bnAbs’ preference for membrane insertion and their rough geometric preferences.

We developed two CG simulation methods to probe the association process. In the first approach termed “spontaneous insertion”, 18 simulation replicates were initiated with the Fab placed in water at different random initial orientations and distances (0.5 – 2 nm) above a pre-assembled HIV-like anionic lipid bilayer, wherein the protein could freely diffuse for 14 µs (Figure 4A). For 4E10, PGZL1, and 10E8, most replicates contacted the membrane and stably inserted for longer than 1 µs; 10, 14, and 12 extended insertion events were captured respectively (Figure 4B-D, Table S2). The simulations also captured both dissociation events and preliminary scanning behaviors, indicating the reversibility and dynamics of the process. As a reference, bovine serum albumin (BSA) was tested similarly. Its documented weak affinity for lipid bilayer association (low millimolar) aligns well with the vastly reduced number of membrane contact events (2/18), minimal overall contact time, and lack of sustained insertion observed (Figure 4E). We also tested 13h11, a non-neutralizing anti-gp41 antibody experimentally shown to have undetectably low lipid bilayer interaction7, 40, as a more relevant control to antibody- based features (Figure 4E). 13h11 exhibited negligible insertion and only sparse short ‘scanning’ events. Thus, these CG simulations readily distinguish MPER bnAbs with high propensity for lipid interaction (low micromolar affinity) from non-specific and low affinity lipid interactions.

Unbiased spontaneous membrane insertion events and semi-biased dissociation events in coarse grain MD simulations

(A) Snapshots of spontaneous insertion event from a Martini model coarse-grain simulation of a 4E10 Fab. The Fab begins in explicit water solvent 1-2 nm above a lipid bilayer, freely diffusing and tumbling in bulk solvent, often resulting in a temporary or permanent insertion event (right).

(B,C,D) 18 replicates of coarse grain Fab systems (4E10, PGZL1, 10E8, respectively), initialized with slightly different Fab orientations relative to lipid bilayer. Frames with Fab contacting the membrane are in green and frames with Fab in water (non-associated) are in white for replicate trajectories of 14 µs each.

(E) 18 replicates of coarse grain BSA (top) or 13h11 (bottom) with different starting orientations relative to the lipid bilayer. Membrane contact (green) or diffusion in water (white) shown over 10 µs time.

(F) Snapshots describing a co-assembling membrane pipeline with 4E10 Fab. An Fab is centered in a box with various rotational orientations in space, explicit water, and lipids randomly arranged within a subset of the box (left). By 30ns, the membrane is fully formed (middle). Fab molecules result in a pre-docked membrane bound conformation and sample a permanent insertion event, intermittent membrane association, or dissociation depending on how the Fab contacts with the membrane (right).

(G, H, I) 40 replicates of 5 µs simulations for coarse grain co-assembling systems (4E10, PGZL1, 10E8, respectively), each with slightly different Fab initial orientations relative to lipid bilayer. Membrane contact is classified as above.

As an alternative strategy to sample bnAbs insertion and favored geometries, we developed a “co-assembly” CG simulation. Each simulation box hosted one Fab at a varied initial orientation amongst a mixture of randomly placed water, ion, and lipid particles as a starting point for subsequent unrestrained MD, leading to spontaneous assembly of the lipid bilayer – often yielding Fab membrane insertion (Figure 4F). This approach increased predisposition for Fab-lipid interactions and reduced wasted time sampling Fab diffusing in bulk water, successfully resulting in more trajectories with Fab inserted in distinct geometries and overall membrane contact time (Figure 4G-I, 5B). However, this method also risks enriching insertion in less favorable states, possibly skewing the conformational distribution. We combated these biases by running more replicates (n=40) of shorter 5 µs simulations. Fab dissociation was also readily observed from initially inserted or ‘scanning’ trajectories, suggesting that coarse-grained models can distinguishing unfavorable and favorable membrane-bound conformations.

Next, we characterized the spectrum of membrane-bound conformations sampled the two CG approaches for each bnAb and compared with our pre-inserted all-atom simulations. We tracked the global Fab domain orientation through the canonical “angle of approach” (angle of the Fab’ long axis relative to the membrane normal vector) alongside a second “rotation angle” tracking Fab rotation around its long axis (Figure 5A). Both CG approaches significantly sampled conformations analogous to the primary state adopted in all-atom trajectories for each bnAb, although with a broader local geometric distribution. Interestingly, PGZL1 and 10E8 heavily sampled alternative membrane-inserted conformations, while 4E10 inserted only in a single conformation. It is unclear whether this observed conformational specificity represents a deeper energy well for 4E10’s primary geometry or is due to the Martini model over-estimating hydrophobic interactions41.

Membrane surface-bound bnAb conformations sampled across multiscale simulations

(A) Graphic of angles defined to describe Fab geometries relative to the normal vector at the membrane’s upper leaflet lipid at the phosphate plane in simulation frames (orange arrow). The canonical “approach angle” defines the long axis of the Fab domain (i.e. the central pseudo-symmetry axis) and membrane normal vector (black arrow). A second “rotational angle” is defines the global domain rotation about the Fab pseudo-symmetry axis relative to the membrane normal vector, based on the short axis traversing the light and heavy chains, which is nearly orthogonal to the Fab’s central axis (red arrow).

(B) Frequency plots of rotation and approach angles from frames of membrane-bound Fabs in MD simulations for 4E10 (blue, top row), PGZL1 (red, middle row), and 10E8 (purple, bottom row). Contour plots depicting frequency maxima for angle pairs sampled are by kernel density estimation. Left column, membrane interaction angles sampled from all-atom simulations with Fabs pre-docked using OPM PPM server prediction. Middle column, geometries from coarse-grain membrane co-assembly simulations. Right column, geometries from unbiased spontaneous insertion coarse-grain simulations. Black dots denote the initial Fab-membrane geometries of starting states for replicate trajectories for each antibody initiated in the lipid bilayer.

Notably, all membrane-associated conformations are mediated by CDR interactions across bnAb systems (Figure S6C-E). Overall, many macroscopic states sampled appear to be biologically relevant, given their geometric similarity with all-atom conformations shown to stably bind phospholipids. For 10E8, spontaneous insertion performed better at sampling geometries akin to those observed in all-atom simulations. Geometries that 4E10 and PGZL1 sampled in co-assembly trajectories overlapped more with all-atom simulations. Nonetheless, we show CG Fab representation can be useful for investigating dynamics of insertion and landscape of surface- bound conformations.

Integrative multi-scale simulations of ab initio formation of bnAb phospholipid complexes

We then tested whether the full process of bnAb association followed by formation of specific phospholipid complexes could be modelled ab initio by integrating spontaneously inserted CG conformations into all-atom trajectories. These simulations assessed compatibility of CG-derived surface geometries to form those specific headgroup interactions, barring minor geometric reorientation. Medoid Fab poses for each of 4E10, PGZl1, and 10E8 were extracted representing 3 structural distinct states after clustering CG simulation frames (subsets A, B, C), then converted to all-atom detail for unbiased MD simulation (Figure 6A-B).

Back-mapping CG membrane-bound geometries to all-atom simulations allows integrative ab initio modeling of the full bnAb insertion process.

(A) Representative frames from membrane-bound 4E10 Fab coarse-grained simulations were back-mapped to all-atom representation to assess the stability and plausibility of those membrane-bound conformations. This coarse-grained-to-all-atom (CG-to-AA) reversion was applied for all medoid frames of interest from each antibody system and used to initiate half-microsecond unbiased all-atom dynamics simulations.

(B) Frequency of membrane interaction angles from coarse grain spontaneous insertion as clustered by geometric substates for 4E10, PGZL1 and 10E8, colored and contoured as in Fig 3B. Corresponding primary all-atom simulations is overlaid (unfilled, black contour frequency density plot).

(C) Conformational geometry sampled upon conversion of CG medoid to an all-atom trajectory. Initial geometry denoted by stars colored matching CG clusters in (B). Frequency and contour plots of conformational angles sampled in stably inserted backmapped all-atom trajectories for 4E10 (left), PGZL1 (middle), and 10E8 (right).

(D) Per-residue interaction profiles for antibody Fab simulations for 4E10 (left), PGZL1 (middle), 10E8 (right) representing each backmapped atomic trajectory, showing CDR-mediated conformations of differing depths and geometries.

(E) Phospholipid headgroup binding and RMSD plots of closest lipid at respective experimentally determined CDR sites for 4E10 (left), PGZL1 (middle), and 10E8 (right), plotted as in Figure 1C or 2B.

For 4E10, trajectories initiated from all three geometries spent extended time (>100 ns) with phospholipid bound at the CDR-H1 site (Figure 6E). These trajectories also drifted back to conformations similar to those in our previous pre-inserted all-atom simulations as measured by the global angles sampled (approach angle, 60- 80°; Figure 6C) and per-residue interaction profile (CDR-H2-H3 insertion, Figure 6D). Back-mapped trajectory A was distinguished by deeper insertion, with the highest >60% CDR-H1 phospholipid binding occupancy (>60%), and the most resemblance to our previous atomic simulations (Figures 1E, 6D). Trajectories initiated from clusters B and C are slightly tilted (rotation angle > 0°) and less extensively inserted, exhibiting lower CDR- H1 phospholipid occupancy (Figure 6C-E). Thus, these integrative 4E10 simulations show how CG simulations could be used to generate stable membrane-bound conformations that align with laboratory experiments and capture the specific phospholipid binding events crucial for bnAbs function – all totally ab initio.

Similarly, for PGZL1, all-atom trajectories starting from CG clusters B and C conformationally converge to similar CDR-H-inserted geometries sampled in previous all-atom MD, and stably bind phospholipid headgroups at CDR-H1 (>50% occupancy) (Figure 6C-E). However, for the trajectory initiation from cluster A, a novel membrane-bound conformation is sampled, which is similarly mediated by CDR-H3 insertion but now supplemented with light-chain contacts. No CDR-H1 lipid complex is detected. This exploration of the additional membrane-bound geometries facilitated by CG sampling highlights alternative kinetically stable states, which may represent modeling artifacts or may capture relevant intermediate conformations aiding membrane association (e.g. preliminary surface-scanning).

10E8 back mapping revealed advantages as well as limitations in the spontaneous insertion CG method. The all-atom trajectory from a CG cluster A initiated from an alternative conformation (approach angle, ∼20°; rotation angle, ∼55°) rapidly dissociated from the membrane within 100 ns. This instability indicates the conformation is likely an artifact, despite being a predominant geometry in CG simulations (Figure 6B,C). Interestingly, the all- atom trajectory starting from CG cluster B trended towards the same novel conformation as cluster A, but remained stably inserted, albeit having shallower CDR-H3 insertion alongside canonical 10E8 CDR-L embedding (Figure 6C-D). Cluster C started and finished in membrane bound orientations most similar to our previous atomistic simulations (Figure 6C), although with a slightly different protein-lipid interaction pattern (Figure 6D). Curiously, the bivalent CDR-L1 FR-L3 phospholipid complex could still be formed with high occupancy for both cluster B and C (>70%, Figure 6E).

We evaluated the possibility of a bivalent IgG interaction with the membrane to provide further mechanistic context with the physiologically relevant molecular structure. Aligning a full-length IgG onto extracted membrane bound Fab conformations suggests it is unlikely that two Fabs simultaneously engage the membrane via low energy conformations due to rotational strain on the IgG hinge region (Figure S7E-F). Together, these integrative modeling data in concert reflect both the value added and the caveats from additional sampling of membrane-bound bnAb conformations from CG trajectories. This ab initio multi-scale simulation approach can identify bnAbs’ apparent lowest energy state with high occupancy CDR-phospholipid complexes. Additionally, sampling a broader range of bnAb surface geometries revealed a wider range of conformations are compatible with forming each bnAb CDR-phospholipid complex, although often with lower occupancy and stability. However, many heavily sampled CG states are likely simulation artifacts and biologically irrelevant, indicated by dissociation or poor insertion patterns upon simulation at atomistic detail. Nonetheless, we demonstrate the utility of this integrative approach for unbiased investigation of the dynamic process of insertion, mapping antibody- membrane interactions, and detecting specific phospholipid binding ab initio at atomic detail.

Atomistic pulling simulations distinguish differences in affinities of experimentally characterized variants

Having established these workflows, we sought to gain a more quantitative energetic view of how these membrane-bound conformations and lipid interactions connect with physiological properties such as affinity and neutralization. Canonical free energy calculations at atomic-detail (e.g. potential of mean force, P.M.F.) requires simulation timescales (10’s of µs)42 prohibitive for characterizing more than a few antibody variants. Here, we developed a more expedient approach, applying constant velocity biased all-atom simulations to determine the rupture force required to dissociate a Fab bound to lipid bilayers. This method provides a binding strength estimate akin to force spectroscopy (Figure 7A). Given that Fabs can sample many lipid interactions and surface conformations, we performed ensemble-based measurements: averaging forces from several replicates of pulling trajectories each initiated with different starting configurations (which were extracted from unbiased MD).

Biased antibody-membrane pulling dissociation simulations approximate Fab-membrane interaction strength for bnAb Fab variants of known lipid binding affinity or neutralization potency.

(A) Schematic of pulling method, a bnAbs Fabs associated with the lipid bilayers is subject to an applied upward dissociation force to the Fab domain center of mass at constant velocity to measure the force required to dissociate the Fab from the bilayer.

(B) The Trp100a-Trp100c motif residues in 4E10 CDR-H3 loop of expected to be lipid-embedded in the bilayer (beige). The double alanine mutant “WAWA” 4E10 variant has experimentally determined significantly reduced affinity to lipid bilayers and lower neutralization potency, due to lack of Trp membrane insertion.

(C) Previously experimentally characterized20 PGZL1 germline-reverted variants, shown as chimera of germline versus matured gene segments, used to approximate antibody properties along its maturation trajectory.

(D) Average force versus distance plots and rupture force (Fmax) calculation for one replicate of pulling wild type 4E10 (WT, blue) and 4E10 WAWA (red) to bias Fab dissociation from the bilayer.

(E) Distribution of rupture forces required for dissociation for different membrane-bound starting conformations for 4E10 (n=9, blue) and 4E10 WAWA (n=11, red), with starting conformations drawn from previous unbiased all-atom simulation ensembles at rest. Outliers as dots

(F) Distributions of rupture forces (n=10) required for membrane dissociation of PGZL1 inferred variants along the maturation pathway, for germline (dark grey, Grm.), intermediate (light grey, Int.), and mature (maroon, Mat.) PGZL1. Outliers as dots

First, to assess suitable pulling velocities, we began dissociating 4E10 from simplified anionic HIV-like lipid bilayers to a fixed distance (1.5 nm) along the membrane normal over different time spans (10, 50, 100, 200 ns), initiating 3 replicates from unique starting conformations (Figure S7A-B). Interestingly, 2 of 3 initial Fab models at their CDR-H3 W-W motif adopt a rotameric state where Trp100a remains inserted in both states but Trp100b projects down towards lipid tails in the “Trp-Down” state while in the other model Trp100b is flipped up towards the lipid-water interface (“Trp-Up” state). As expected, rupture forces decrease with increasing simulation time (slower pulling velocity), plateauing between 100 and 200 ns (Figure S7A). Surprisingly, the rupture force was nearly identical (< 2% difference) between the two Trp-Down pulling trajectories at both 50 ns and 100 ns (78.5 and 56.2 kJ/mol·nm2, respectively), and was consistently higher than forces from the Trp-Up conformation (27 and 47% greater, respectively) (Figure S7A). This cursory trial of bnAb dissociation force calculations exhibited promising conformational sensitivity in distinguishing different CDR loop protein-lipid interactions and remarkable precision between similar starting Fab conformations.

We next benchmarked this technique by comparing behavior of 4E10 with its well-studied WAWA variant (W100aA-W100bA CDR-H3 mutant) (Figure 7B). WAWA fragments have similar Kd to MPER antigens, but significantly reduced affinity to HIV-like empty liposomes and drastically reduced neutralization efficacy (>100 fold higher IC50)21, 25, 28, 43. Thus, we hypothesized that in silico estimation could assist in making the molecular link between WAWA’s reduced membrane binding and defective neutralization. Across replicates from different starting membrane-inserted conformations for 4E10 (n=9) and WAWA (n=11), rupture force calculations showed average forces (± S.E.M.) of 63.8 ± 2.8 kJ/mol·nm2 and 48.0 ± 2.6 kJ/mol·nm2 respectively (Figures 7D-E, S7D), correctly indicating the 4E10 ensemble has a much stronger membrane interaction (32 ± 13 %; p-value < 0.001). Thus, this ensemble analysis approach robustly discerned antibodies of high (∼1 μM) and low membrane affinity previous measured experimentally and should prove useful in investigating the apparent correlations between bnAb molecular features, lipid bilayer interactions, and neutralization potency. Notably, each simulation set required 30-40 hours (replicates in parallel) in a modern computing environment, which is less than 10% needed for a PMF42 and reflects medium throughout scalable to hundreds of antibodies per year.

Finally, we evaluated the membrane interactions of experimentally characterized PGZL1 variants that recapitulate possible stages of a germline maturation pathway20 (Figure 7C). Mature PGZL1 has relatively high affinity to the epitope peptide (Kd = 10 μM) and demonstrates great breadth and potency, neutralizing 84% of a 130 strain panel. An “Intermediate” PGZL1 with V-gene germline reversion (gVmDmJ, having CDR-H3 100% sequence identity, Figure S7D) showed a modest reduction in affinity (Kd = 64 μM), but greatly reduced neutralization potency and breadth (12% of 130 strains)20. The loss of neutralization efficacy is far greater than expected from the modestly lowered antigen affinity alone, and was hypothesized to be due to vastly reduced propensity for lipid interactions of this variant20. The completely reverted “Germline” version (gVgDgM) had no detectable MPER affinity or neutralization. To predict the initial membrane-docked Germline and Intermediate Fab models, we performed the aforementioned unrestrained all-atom workflow starting with hydrophobicity- optimized inserted conformations. The PGZL1 Fabs all had similar predicted insertion geometries, facilitated by long and relatively hydrophobic CDR-H3 conserved across variants (Figure S7D). CDR-H1-phospholipid complexes were not formed during Intermediate or Germline unbiased simulation.

Ten unique lipid-inserted conformations were extracted from the MD ensemble to initiate replicates of biased membrane dissociation simulations. Significantly lower forces were required to dissociate Intermediate and Germline PGZL1 from bilayers, 37.5 ± 14.9 kJ/mol·nm2 and 41.2 ± 11.7 kJ/mol·nm2 respectively, than for mature PGZL1, 56.1 ± 11.6 kJ/mol·nm2 (p < 0.003) (Figure 7E). The Germline PGZL1 likely represents a kinetically stable but low affinity interaction, setting the low baseline of rupture forces required to dissociate CDR-inserted antibodies. While the difference in Germline and Intermediate force distributions was not significantly different (p > 0.4), more higher force events were observed for the Intermediate variant versus Germline. Thus, PGZL1 more favorably interacts with anionic HIV-like membranes than inferred maturation pathway precursors. We did not disentangle the energetic contributions to membrane affinity of the CDR-H1 phospholipid complex from other protein-lipid interactions determining mature PGZL1’s enhanced membrane interaction here, but the roles of these features should be explored in future work. These results provide evidence that phospholipid binding features are acquired and honed along a bnAb maturation pathway, entrenching the molecular link between positive selection of bnAb lipid binding properties and acquiring neutralization efficacy.

Discussion

Here, we present a roadmap for applying integrative molecular simulations to characterize the biophysical underpinning of lipid membrane interactions within the mechanism of HIV MPER bnAbs. The methods developed and principles extracted improve understanding of maturation pathways for antibodies targeting membrane-proximal epitopes and should enrich data-driven design of HIV immunogens. Our detailed simulations supplement in vitro binding and crystallographic evidence in demonstrating that bnAbs develop highly specific phospholipid interactions that facilitate access to the MPER epitope and can participate in epitope-paratope interface in context of full lipid bilayers. Furthermore, the modeling approaches demonstrated here could be used both retrospectively or proactively: to identify, characterize, and even engineer membrane-targeting elements in antibodies.

Further, these simulations can serve future examination of molecular details concerning the genetic origins and developmental pathways for incipient lipid-binding antibodies. Through natural or vaccine-induced immunity, coaxing the immune system to develop both lipid and antigen affinity (possibly even cooperativity) during maturation is a difficult task, and hindered by downregulation of membrane binding precursors1, 3, 6, 11. Likewise, MPER bnAbs are difficult to induce and have only been isolated from patients with chronic HIV- infection and sustained immunosuppression17, 18, 35, 43, 44. Modern vaccine strategies often elicit certain subsets of precursor B cells, often targeting specific germline genes, in attempt to guide antibodies’ mature of specific molecular features intended for the targeted epitopes45, 46. The question remains whether particular germline genes and lineages are privileged for successful maturation as membrane-targeting (or MPER-targeting) antibodies. This notion is supported by 4E10 and PGZL1, which share a germline gene (VH1-69)20 and demonstrate near identical membrane interactions. To this end, germlines genes with inherent basal membrane affinity may be favored as MPER bnAb precursors, potentially circumventing auto-immune checkpoints during maturation or presenting unique scaffolds compatible with further evolution of lipid binding properties. Specific CDR lipid-binding motifs are predicted to be encoded early in maturation processes for PGZL120, namely the CDR-H1 loop (Figure S7D). These features are complemented by mutated membrane contacting residues in the FR-H3 surface and CDR-H3 loop, which are incorporated later in the maturation process (Figure S7D). Across MPER bnAbs, membrane-binding features are likely acquired at strategic timepoints throughput development to balance poly- specificity and evade autoimmune checkpoints47. Experimental databases of germlines genes usage and of B-cell repertoires sequenced during immunization courses would be ideally paired with the simulations described here to investigate in vivo filtering of lineages with lipid-binding variants and assess possible rules of the autoimmune system48.

Beyond gp41-targeting antibodies and lipid antibodies in autoinflammatory diseases49, 50 or microbial infections51, 52, we suspect a broader scope of positive outcomes from antibody tolerance and interaction with lipid bilayers may occur in Nature and is currently underestimated and underutilized. Although host cross-specificity of 4E10 is well documented, the vastly reduced poly-reactivity of PGZL1 and 10E8 inspires optimism that maturation of phospholipid interactions and membrane tolerance may be viable to incorporate into emerging vaccine design strategies and therapeutic antibodies for other membrane proteins to facilitate access of conserved, buries epitopes33. Currently, the breadth of analogous antibodies targeting membrane protein while making extensive or cooperative lipid interactions is poorly explored. The simulation procedures and principles described here are well positioned to investigate this outstanding question and could help define broader chemical rules for design of membrane-interfacing antibodies.

Supporting information

eLife Assessment

This valuable study reports multi-scale molecular dynamics simulations to investigate a class of highly potent antibodies that simultaneously engage with the HIV-1 Envelope trimer and the viral membrane. The work provides insights into how broadly neutralizing antibodies associate with lipids proximal to membrane-associated epitopes to drive neutralization. However, the evidence for rules for lipid recognition in antibodies is still incomplete. In addition, the text would benefit from clearer subsections that delineate discrete mechanistic discoveries.

Reviewer #1 (Public Review):

Previous experimental studies demonstrated that membrane association drives avidity for several potent broadly HIV-neutralizing antibodies and its loss dramatically reduces neutralization. In this study, the authors present a tour de force analysis of molecular dynamics (MD) simulations that demonstrate how several HIV-neutralizing membrane-proximal external region (MPER)-targeting antibodies associate with a model lipid bilayer.

First, the authors compared how three MPER antibodies, 4E10, PGZL1, and 10E8, associated with model membranes, constructed with a lipid composition similar to the native virion. They found that the related antibodies 4E10 and PGZL1 strongly associate with a phospholipid near heavy chain loop 1, consistent with prior crystallographic studies. They also discovered that a previously unappreciated framework region between loops 2-3 in the 4E10/PGZL1 heavy chain contributes to membrane association. Simulations of 10E8, an antibody from a different lineage, revealed several differences from published X-ray structures. Namely, a phosphatidylcholine binding site was offset and includes significant interaction with a nearby framework region.

Next, the authors simulate another MPER-targeting antibody, LN01, with a model HIV membrane either containing or missing an MPER antigen fragment within. Of note, LN01 inserts more deeply into the membrane when the MPER antigen is present, supporting an energy balance between the lowest energy conformations of LN01, MPER, and the complex. Additional contacts and conformational restraints imposed by ectodomain regions of the envelope glycoprotein, however, remain unaddressed-the size of such simulations likely runs into technical limitations including sampling and compute time.

The authors next established course-grained (CG) MD simulations of the various antibodies with model membranes to study membrane embedding. These simulations facilitated greater sampling of different initial antibody geometries relative to membrane. Distinct geometries derived from CG simulations were then used to initialize all-atom MD simulations to study insertion in finer detail (e.g., phospholipid association), which largely recapitulate their earlier results, albeit with more unbiased sampling. The multiscale model of an initial CG study with broad geometric sampling, followed by all-atom MD, provides a generalized framework for such simulations.

Finally, the authors construct velocity pulling simulations to estimate the energetics of antibody membrane embedding. Using the multiscale modelling workflow to achieve greater geometric sampling, they demonstrate that their model reliably predicts lower association energetics for known mutations in 4E10 that disrupt lipid binding. However, the model does have limitations: namely, its ability to predict more subtle changes along a lineage-intermediate mutations that reduce lipid binding are indistinguishable from mutations that completely ablate lipid association. Thus, while large/binary differences in lipid affinity might be predictable, the use of this method as a generative model are likely more limited.

The MD simulations conducted throughout are rigorous and the analysis are extensive. However, given the large amount of data presented within the manuscript, the text would benefit from clearer subsections that delineate discrete mechanistic discoveries, particularly for experimentalists interested in antibody discovery and design. One area the paper does not address involves the polyreactivity associated with membrane binding antibodies-MD simulations and/or pulling velocity experiments with model membranes of different compositions, with and without model antigens, would be needed. Finally, given the challenges in initializing these simulations and their limitations, the text regarding their generalized use for discovery, rather than mechanism, could be toned down.

Overall, these analyses provide an important mechanistic characterization of how broadly neutralizing antibodies associate with lipids proximal to membrane-associated epitopes to drive neutralization.

Reviewer #2 (Public Review):

In this study, Maillie et al. have carried out a set of multiscale molecular dynamics simulations to investigate the interactions between the viral membrane and four broadly neutralizing antibodies that target the membrane proximal exposed region (MPER) of the HIV-1 envelope trimer. The simulation recapitulated in several cases the binding sites of lipid head groups that were observed experimentally by X-ray crystallography, as well as some new binding sites. These binding sites were further validated using a structural bioinformatics approach. Finally, steered molecular dynamics was used to measure the binding strength between the membrane and variants of the 4E10 and PGZL1 antibodies.

The conclusions from the paper are mostly well supported by the simulations, however, they remain very descriptive and the key findings should be better described and validated. In particular:

It has been shown that the lipid composition of HIV membrane is rich in cholesterol [1], which accounts for almost 50% molar ratio. The authors use a very different composition and should therefore provide a reference. It has been shown for 4E10 that the change in lipid composition affects dynamics of the binding. The robustness of the results to changes of the lipid composition should also be reported.

The real advantage of the multiscale approach (coarse grained (CG) simulation followed by a back-mapped all atom simulation) remains unclear. In most cases, the binding mode in the CG simulations seem to be an artifact.

The results reported in this study should be better compared to available experimental data. For example how does the approach angle compare to cryo-EM structure of the bnAbs engaging with the MPER region, e.g. [2-3]? How do these results from this study compare to previous molecular dynamics studies, e.g.[4-5]?

References

1. Brügger, Britta, et al. "The HIV lipidome: a raft with an unusual composition." Proceedings of the National Academy of Sciences 103.8 (2006): 2641-2646.

2. Rantalainen, Kimmo, et al. "HIV-1 envelope and MPER antibody structures in lipid assemblies." Cell Reports 31.4 (2020).

3. Yang, Shuang, et al. "Dynamic HIV-1 spike motion creates vulnerability for its membrane-bound tripod to antibody attack." Nature Communications 13.1 (2022): 6393.

4. Carravilla, Pablo, et al. "The bilayer collective properties govern the interaction of an HIV-1 antibody with the viral membrane." Biophysical Journal 118.1 (2020): 44-56.

5. Pinto, Dora, et al. "Structural basis for broad HIV-1 neutralization by the MPER-specific human broadly neutralizing antibody LN01." Cell host & microbe 26.5 (2019): 623-637.

Author response:

The following is the authors’ response to the current reviews.

Reviewer #1 (Public review):

Previous experimental studies demonstrated that membrane association drives avidity for several potent broadly HIV-neutralizing antibodies and its loss dramatically reduces neutralization. In this study, the authors present a tour de force analysis of molecular dynamics (MD) simulations that demonstrate how several HIV-neutralizing membrane-proximal external region (MPER)-targeting antibodies associate with a model lipid bilayer.

First, the authors compared how three MPER antibodies, 4E10, PGZL1, and 10E8, associated with model membranes, constructed with two lipid compositions similar to native viral membranes. They found that the related antibodies 4E10 and PGZL1 strongly associate with a phospholipid near heavy chain loop 1, consistent with prior crystallographic studies. They also discovered that a previously unappreciated framework region between loops 2-3 in the 4E10/PGZL1 heavy chain contributes to membrane association. Simulations of 10E8, an antibody from a different lineage, revealed several differences from published X-ray structures. Namely, a phosphatidylcholine binding site was offset and includes significant interaction with a nearby framework region. The revised manuscript demonstrates that these lipid interactions are robust to alterations in membrane composition and rigidity. However, it does not address the reverse-that phospholipids known experimentally not to associate with these antibodies (if any such lipids exist) also fail to interact in MD simulations.

Next, the authors simulate another MPER-targeting antibody, LN01, with a model HIV membrane either containing or missing an MPER antigen fragment within. Of note, LN01 inserts more deeply into the membrane when the MPER antigen is present, supporting an energy balance between the lowest energy conformations of LN01, MPER, and the complex. These simulations recapitulate lipid binding interactions solved in published crystallographic studies but also lead to the discovery of a novel lipid binding site the authors term the "Loading Site", which could guide future experiments with this antibody.

The authors next established course-grained (CG) MD simulations of the various antibodies with model membranes to study membrane embedding. These simulations facilitated greater sampling of different initial antibody geometries relative to membrane. These CG simulations , which cannot resolve atomistic interactions, are nonetheless compelling because negative controls (ab 13h11, BSA) that should not associate with membrane indeed sample significantly less membrane.

Distinct geometries derived from CG simulations were then used to initialize all-atom MD simulations to study insertion in finer detail (e.g., phospholipid association), which largely recapitulate their earlier results, albeit with more unbiased sampling. The multiscale model of an initial CG study with broad geometric sampling, followed by all-atom MD, provides a generalized framework for such simulations.

Finally, the authors construct velocity pulling simulations to estimate the energetics of antibody membrane embedding. Using the multiscale modelling workflow to achieve greater geometric sampling, they demonstrate that their model reliably predicts lower association energetics for known mutations in 4E10 that disrupt lipid binding. However, the model does have limitations: namely, its ability to predict more subtle changes along a lineage-intermediate mutations that reduce lipid binding are indistinguishable from mutations that completely ablate lipid association. Thus, while large/binary differences in lipid affinity might be predictable, the use of this method as a generative model are likely more limited.

The MD simulations conducted throughout are rigorous and the analysis are extensive, creative, and biologically inspired. Overall, these analyses provide an important mechanistic characterization of how broadly neutralizing antibodies associate with lipids proximal to membrane-associated epitopes to drive neutralization.

Reviewer #2 (Public review):

In this study, Maillie et al. have carried out a set of multiscale molecular dynamics simulations to investigate the interactions between the viral membrane and four broadly neutralizing antibodies that target the membrane proximal exposed region (MPER) of the HIV-1 envelope trimer. The simulation recapitulated in several cases the binding sites of lipid head groups that were observed experimentally by X-ray crystallography, as well as some new binding sites. These binding sites were further validated using a structural bioinformatics approach. Finally, steered molecular dynamics was used to measure the binding strength between the membrane and variants of the 4E10 and PGZL1 antibodies.

The use of multiscale MD simulations allows for a detailed exploration of the system at different time and length scales. The combination of MD simulations and structural bioinformatics provides a comprehensive approach to validate the identified binding sites. Finally, the steered MD simulations offer quantitative insights into the binding strength between the membrane and bnAbs.

While the simulations and analyses provide qualitative insights into the binding interactions, they do not offer a quantitative assessment of energetics. The coarse-grained simulations exhibit artifacts and thus require careful analysis.

This study contributes to a deeper understanding of the molecular mechanisms underlying bnAb recognition of the HIV-1 envelope. The insights gained from this work could inform the design of more potent and broadly neutralizing antibodies.

Recommendations for the authors:

Reviewing Editor:

We recommend the authors remove the figure and section related to bnAb LN01, perform additional analysis (e.g., further expanding on the differences in antibody binding in the presence or absence of antigen), and present this as a separate manuscript in a follow-up study.

We consider the analysis of a bnAb with a transmembrane antigen and of LN01 as essential to the manuscript and novel results. Study of LN01 provides many insights unique from the other MPER bnAbs in this study. We agree further characterization of LN01 and bnAbs with transmembrane antigen or full-length Env are intriguing and necessary to complete the full mechanistic understanding of lipid-associated antibodies. LN01 section in this paper is novel in the field and demonstrates the preliminary evidence motivating further work, which we agree are beyond the scope of this already long detailed study.

Reviewer #1 (Recommendations for the authors):

I appreciate the degree to which the authors responded to my previous points raised in the private review, including edits where I might have missed something in the manuscript or relevant literature. I imagine such a point-by-point response was quite onerous. Thank you also for balancing presentation/clarity with content/rigor considering the large information content of this manuscript; in silico results are inherently hard to present given the delicate balance between rigorous validation and novel information content. I apologize if I repeat points raised and addressed previously and commend the authors on their revised study, which is much improved in clarity; any additional revisions are of course entirely at your discretion.

"...now having more diversity in lipid headgroup chemistries" references the wrong figure-it should be: Figure 2-figure supplement 2A-C. The incorrect figure is also referenced again several sentences down: "...relevant CDR and framework surface loops..."

Thank you for pointing out this error. We have corrected figure references.

"One shared conformational difference observed for these bnAbs the higher cholesterol bilayers was slightly more extensive and broader interaction profiles as well as modestly deeper embedding of the relevant CDR and framework surfaces loops" please rephrase

Thank you for this suggestion. We rephrased this for improved clarity and flow.

"These results bolster the feasibility for using all-atom MD as an in silico platform to explore differential phospholipid affinity at these sites (i.e., specificity studies) and influence on antibody preferred conformation as membrane composition and lipid chemistry are systematically varied" Please tone down these speculations-you have demonstrated that simulations are robust to different headgroup chemistries but have not provided evidence for the exclusion of lipids that are known not to associate with these antibodies.

We rephrased this speculation to highlight the potential of this application. We also emphasize future studies that would be required to achieve this application in the following sentence.

“These results motivate use of all-atom MD as an in silico approach for exploring differential phospholipid affinity at these sites…”

Figure 2A: Specify which PDB entry corresponds to the displayed crystal structures in the main figure or caption.

We clarified these PDB entries in the figure caption.

Check reference formatting in supplemental figures when generating VOR.

I am not sure how relevant this might be to the claims of Figure 2-figure supplement 3, but AlphaFold3-based phospholigand docking might provide an additional orthogonal approach if relevant ligand(s) are available for such analysis (particularly for the newly proposed 10E8 POPC complex).

Thank you for this suggestion. AI/ML based prediction methods like AF3 and RoseTTAFold All-Atom (RFAA) are interesting new methods that have come since our initial submission. We’ve decided these experiments are beyond the scope of this already long and detailed study. We have added a sentence suggesting use of these methods in future work.

"We next studied bnAb LN01 to interrogate differences" --> this transition still reads a bit unclear. Why shift gears and change antibodies? Also, while you do go into its interactions both +/- antigen, there's no lead into the simulation initialization with and without antigen to guide the reader into the comparisons you will draw in the figure. Also, the order of information presentation is a bit strange, where the rationale for choosing a single monomeric helix is brought up in the middle of the paragraph instead of at the beginning of the section. In the next paragraph, it goes back to the initialization of the membrane composition again, which feels a bit disorganized-I do appreciate the unique challenge of having to weave through so much quality data! In fact, if you were to conduct simulations of membrane + antigen vs. membrane + LN01 vs. membrane + LN01 + antigen, I am tempted to say that this could be removed from this manuscript and flow better as a paper in and of itself.

We thank the reviewer for the suggestion to improve the writing style. We feel this section adds a lot of value to the manuscript, so we will keep it in the paper and improved the transition as well as rationale.

We selected to study the additional antibody LN01 and the monomeric MPER-TM antigen conformation because of the existing structural evidence available without additional creative model building. This rationale has been updated in the new text.

We changd the order of information as suggested, moving the rationale for antigen fragment earlier in the paragraph followed by the background of the lipids sites from the crystal that can lead into simulation set-up. We clarified the simulation initialization was similar for systems with and without antigen in the opening sentence of the paragraph

"previously observed snorkeling and hydration of TM Arg686" --> Is this R696 (numbering could be different based on the particular Env)?

Thank you for noting this typo, we have corrected the numbering.

Potential font color issue with Figure 3-Figure supplement 1 B and part of A text-could be fixed in typesetting.

The discussion reads very well. Is it possible to direct antibody maturation, even in an engineered context, towards membrane affinity without increasing immunogenic polyreactivity? This is mentioned very briefly and cited with ref 36, but I would be interested in the author's thoughts on this topic.

We thank the reviewer for the insightful idea to explore in future work. Our conclusion alludes to possibly artificially evolving membrane affinity studied by MD, as done in vitro by Nieva and co-workers. Because the hypothetical nature, we’ve chosen not to elaborate on those ideas from this manuscript.

Reviewer #2 (Recommendations for the authors):

To ensure reproducibility and facilitate further research, the authors should publicly deposit the code for running the MD simulations and analyses (e.g., on GitHub) along with the underlying data used in the study (e.g., on Zenodo.org).

We appreciate the consideration for open-source code and analysis. Representative code and simulation trajectories were uploaded to the following repositories:

https://github.com/cmaillie98/mper_bnAbs.git

https://zenodo.org/records/13830877

—-

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public Review):

Previous experimental studies demonstrated that membrane association drives avidity for several potent broadly HIV-neutralizing antibodies and its loss dramatically reduces neutralization. In this study, the authors present a tour de force analysis of molecular dynamics (MD) simulations that demonstrate how several HIV-neutralizing membrane-proximal external region (MPER)-targeting antibodies associate with a model lipid bilayer.

First, the authors compared how three MPER antibodies, 4E10, PGZL1, and 10E8, associated with model membranes, constructed with a lipid composition similar to the native virion. They found that the related antibodies 4E10 and PGZL1 strongly associate with a phospholipid near heavy chain loop 1, consistent with prior crystallographic studies. They also discovered that a previously unappreciated framework region between loops 2-3 in the 4E10/PGZL1 heavy chain contributes to membrane association. Simulations of 10E8, an antibody from a different lineage, revealed several differences from published X-ray structures. Namely, a phosphatidylcholine binding site was offset and includes significant interaction with a nearby framework region.

Next, the authors simulate another MPER-targeting antibody, LN01, with a model HIV membrane either containing or missing an MPER antigen fragment within. Of note, LN01 inserts more deeply into the membrane when the MPER antigen is present, supporting an energy balance between the lowest energy conformations of LN01, MPER, and the complex. Additional contacts and conformational restraints imposed by ectodomain regions of the envelope glycoprotein, however, remain unaddressed-the size of such simulations likely runs into technical limitations including sampling and compute time.

The authors next established course-grained (CG) MD simulations of the various antibodies with model membranes to study membrane embedding. These simulations facilitated greater sampling of different initial antibody geometries relative to membrane. Distinct geometries derived from CG simulations were then used to initialize all-atom MD simulations to study insertion in finer detail (e.g., phospholipid association), which largely recapitulate their earlier results, albeit with more unbiased sampling. The multiscale model of an initial CG study with broad geometric sampling, followed by all-atom MD, provides a generalized framework for such simulations.

Finally, the authors construct velocity pulling simulations to estimate the energetics of antibody membrane embedding. Using the multiscale modelling workflow to achieve greater geometric sampling, they demonstrate that their model reliably predicts lower association energetics for known mutations in 4E10 that disrupt lipid binding. However, the model does have limitations: namely, its ability to predict more subtle changes along a lineage-intermediate mutations that reduce lipid binding are indistinguishable from mutations that completely ablate lipid association. Thus, while large/binary differences in lipid affinity might be predictable, the use of this method as a generative model are likely more limited.

The MD simulations conducted throughout are rigorous and the analysis are extensive. However, given the large amount of data presented within the manuscript, the text would benefit from clearer subsections that delineate discrete mechanistic discoveries, particularly for experimentalists interested in antibody discovery and design. One area the paper does not address involves the polyreactivity associated with membrane binding antibodies-MD simulations and/or pulling velocity experiments with model membranes of different compositions, with and without model antigens, would be needed. Finally, given the challenges in initializing these simulations and their limitations, the text regarding their generalized use for discovery, rather than mechanism, could be toned down.

Overall, these analyses provide an important mechanistic characterization of how broadly neutralizing antibodies associate with lipids proximal to membrane-associated epitopes to drive neutralization.

Reviewer #2 (Public Review):

In this study, Maillie et al. have carried out a set of multiscale molecular dynamics simulations to investigate the interactions between the viral membrane and four broadly neutralizing antibodies that target the membrane proximal exposed region (MPER) of the HIV-1 envelope trimer. The simulation recapitulated in several cases the binding sites of lipid head groups that were observed experimentally by X-ray crystallography, as well as some new binding sites. These binding sites were further validated using a structural bioinformatics approach. Finally, steered molecular dynamics was used to measure the binding strength between the membrane and variants of the 4E10 and PGZL1 antibodies.

The conclusions from the paper are mostly well supported by the simulations, however, they remain very descriptive and the key findings should be better described and validated. In particular:

It has been shown that the lipid composition of HIV membrane is rich in cholesterol [1], which accounts for almost 50% molar ratio. The authors use a very different composition and should therefore provide a reference. It has been shown for 4E10 that the change in lipid composition affects dynamics of the binding. The robustness of the results to changes of the lipid composition should also be reported.

The real advantage of the multiscale approach (coarse grained (CG) simulation followed by a back-mapped all atom simulation) remains unclear. In most cases, the binding mode in the CG simulations seem to be an artifact.

The results reported in this study should be better compared to available experimental data. For example how does the approach angle compare to cryo-EM structure of the bnAbs engaging with the MPER region, e.g. [2-3]? How do these results from this study compare to previous molecular dynamics studies, e.g.[4-5]?

References

(1) Brügger, Britta, et al. "The HIV lipidome: a raft with an unusual composition." Proceedings of the National Academy of Sciences 103.8 (2006): 2641-2646.

(2) Rantalainen, Kimmo, et al. "HIV-1 envelope and MPER antibody structures in lipid assemblies." Cell Reports 31.4 (2020).

(3) Yang, Shuang, et al. "Dynamic HIV-1 spike motion creates vulnerability for its membrane-bound tripod to antibody attack." Nature Communications 13.1 (2022): 6393.

(4) Carravilla, Pablo, et al. "The bilayer collective properties govern the interaction of an HIV-1 antibody with the viral membrane." Biophysical Journal 118.1 (2020): 44-56.

(5) Pinto, Dora, et al. "Structural basis for broad HIV-1 neutralization by the MPER-specific human broadly neutralizing antibody LN01." Cell host & microbe 26.5 (2019): 623-637.

Considering reviewer suggestions, we slightly reorganized the results section into specific sub-sections with headings and changed the order in which key results were presented to allow the subsequent analysis more accessible for readers. Supplemental materials were redistributed into eLife format, having each supplemental item grouped to a corresponding main figure. Many slightly detail modifications were made to figures (mostly supplemental items) without changing their character, such as clearer axes labels or revised annotations within panels.

The major additions within the results sections based on the reviews were:

(1) An expanded the comparison between our simulation analyses to previous simulations and to existing cryo-EM structural evidence for MPER antibodies’ membrane orientation the context of full-length antigen, resulting in new supplemental figure panels.

(2) New atomistic simulations of 10E8, PGZL1, and 4E10 evaluating the phospholipid binding predictions in a different lipid composition more closely modeling HIV membranes.

Minor edits to the analyses and interpretations include:

(1) Outlining the geometric components contributing to variance in substates after clustering the atomistic 10E8, 4E10, and PGZL1 simulations.

(2) Better defining the variance and durability of membrane interactions within and across systems in the coarse grain methods section.

(3) Removed interpretations in the original results sections regarding polyreactivity and energetics for MPER bnAbs that were not explicitly supported by data.

(4) More context of the prevenance of bnAb loop geometries in structural informatics section

(5) Rationale for the choice of the continuous helix MPER-TM conformation in LN01-antigen conformations, and citations to previous gp41 TM simulations.

(6) Removed language on the novelty of the coarse grain and steered pulling simulations as newly developed approaches; tempering the potential discriminating power and applications of those approaches, in light of their limitations.

The discussion was revised to provide more novel context of the results within the field, including discussing direct relevance of the simulation methods for evaluating immune tolerance mechanisms and into antibody engineering. We have shared custom scripts used for molecular dynamics analysis on github (https://github.com/cmaillie98/mper_bnAbs.git) and uploaded trajectories to a public repository hosted on Zenodo (https://zenodo.org/records/13830877).

Recommendations for the authors:

Below, I provide an extensive list of minor edits associated with the text and figures for the authors to consider. I provide these with the hope of increasing the accessibility of the manuscript to broader audiences but leave changes to the discretion of the authors.

Text/clarity

Figure 1 main text

The main text discussing Figure 1 is disorganized, making the analysis difficult to follow. I would suggest the following: moving the sentence, "4E10 and PG2L1 are structurally homologous" immediately after the paragraph discussing the simulation initiation. Then, add a sentence that directly compares their experimental affinity, neutralization, and polyreactivity of 4E10 and PG2L1 (later, an unintroduced idea pops up, "These patterns may in part explain 4E10's greater polyreactivity"). Next, lead into the discussion of the MD simulation data with something to the effect of: "Given these similarities, we first compared mechanisms of membrane insertion between 4E10 and PG2L1 to bolster confidence in our predictions". Later, the sentence "Across 4E10 and PGZL1 simulations, the bound lipid phosphates"

We thank the reviewer for the suggestion and we have restructured the beginning of the results to implement this style: to first introduce then discuss the comparative PGZL1 & 4E10 results, i.e. Figure 1 plus associated supplements.

In the background and the introduction text leading up to Figure 1, CDR-H3 is discussed at length, however, the first figure focuses almost entirely on how CDR-H1 coordinates a lipid phosphate headgroup. Are there experimental mutations in this loop that do not affect affinity (e.g., to a soluble gp41 peptide), but do affect neutralization (like the WAWA mutation for CDR-H3, discussed later)?

We have altered the Introduction (para 2) and Results (4E10/PGZL1 sub-section) to give more balanced discussion of CDRs H1 & H3. That includes referencing experimental data addressing the reviewer’s question; a PGZL1 clone H4K3 where mutations to CDRH1 were introduced and shown have minimal impact on affinity to MPER peptide via ELISA and BLI, but those mutant bnAbs had significantly reduced neutralization efficacy (PMC6879610).

The sentence "These phospholipid binding events were highly stable, typically persisting for hundreds of nanoseconds" should be moved down to immediately precede, "[However], in a PGZL1 simulation, we observed a". This would be a good place for a paragraph break following, "Thus, these bnABs constitutively", since this block of text is very long.

Similarly, the sentence and parts of the section, "Likewise, the interactions coordinating the lipid phosphate oxygens at CDR-H1" more appropriately belongs immediately before or after the sentence, "Our simulations uncover the CDR-lipid interactions that are the most feasible".

Thank you for the detailed guidance in reorganizing the Figure 1 results. We followed the advice to directly compare 4E10 and PGZL1 results separately from 10E8, moving those sections of text appropriately. New paragraph breaks were added to improve accessibility and flow of concepts throughout the Results.

In the sentence, "our simulations uncover CDR-lipid interactions that are the most feasible and biologically relevant in the context of a full [HIV] lipid bilayer... validation to which of the many possible ions" à have you confidently determined lipid binding and positioning outside of the site validated in figure 1? Which site(s) are these referencing? The next two sentences then introduce two new ideas on the loop backbone stability then lead into lipid exchange, which is a bit jarring.

We have adjusted the language concerning the putative ions/lipids electron density across the many PGZL1 and 4E10 crystal structures, and additionally make the explicit point that we confidently determined the lack of lipid binding outside of the site focused on in Figure 1.

“… both bnAbs showed strong hotspots for a lipid phosphate bound within the CDR-H1 loops, with minimal phospholipid or cholesterol ordering around the proteins elsewhere. The simulated lipid phosphates bound within CDR-H1 have exceptional overlap with electron densities and atomic details of modelled headgroups from respective lipid-soaked co-crystal structures…”

Figure 2 main text

"We similarly investigated bnAb 10E8" - Please make this a separate subheader, the block text is very long up to this point.

Thank you for the suggestion. We introduced a sub-header to separate work on 10E8 all-atom simulations.

"we observed a POPC complexed with... modelled as headgroup phosphoglycerol anions..." - please cite the references within the text.

Thank you for pointing out this missing reference, we added the appropriate reference.

"One striking and novel observation" - please remove the phrase "striking" throughout, for following best practices in scientific writing (PMC10212555)-this is generally well-done throughout.

We removed “striking” from our text per your suggestion.

"This CDR-L1 site highlights... (>500 fold) across HIV strains" - How much do R29 and Y32 also contribute to antigen binding and the conformation of this loop? These mutants also decreased Kd by approximately 20X, and based on the co-crystal structure with the TM antigen (PDB: 4XCC), seem to play a more direct role in antigen contact. Additionally, these residues should be highlighted on a figure, otherwise it's difficult to understand why they are important for membrane association.

We thank the reviewer for deep engagement to these supporting experimental details. The R29A+Y32A 10E8 mutant referenced in the text showed only 4-fold Kd increase, a modest change for an SPR binding experiment. Whereas R29E+Y32E 10E8 mutant resulted in 40x Kd increase, the “20x” the reviewer refers to. Both 10E8 mutants showed similar drastically reduced breadth and potency of over 2 orders of magnitude on average.

These mutated CDR-L1 residues are not directly involved in antigen contact and adopt the same loop helix conformation when antigen is bound. A minor impact on antigen binding affinity could be due altering pre-organization of CDR loops upon losing interactions from the Tyr & Arg sidechains - particularly Tyr31 in contact with CDR-H3.

As per the suggestion, clearer annotated figure panel denoting these sidechains has been added to Figure 2-Figure Supplement 1 for 10E8 analysis.

"Structural searches querying... identified between 10^5 and 2*10^6..." - why is this value represented as such a large range? Does this depend on the parameters used for analysis? Please clarify.

Additionally, how prevalent are any random loop conformations compared to the ones you searched? It's otherwise difficult to attribute number of occurrences within the 2 A cutoff to biological significance, as this number is not put in context.

We appreciate the reviewers comment to contextualize the range and relative frequency of the bnAb loop conformations. RMSD and length of loop are the key parameters, which can be controlled by searching reference loops of similar length. The main point of the backbone-level searching is simply to imply the bnAb loops are not particularly rare when comparing loops of similar length.

We did as was suggested and added comparison to random loops of the same length to the main text, including a new Supplementary Table 4.

“…identified between 105 to 2∙106 geometrically similar sub-segments within natural proteins (<2 Å RMSD)40, reflecting they are relatively prevalent (not rare) in the protein universe, comparing well with frequency of other surface loops of similar length in antibodies (Supplementary Table 3).”

"We next examined the geometries" could start after its own new subheading. Moreover, while there's an emphasis on tilt for neutralization, there is not a figure clearly modelling the proposed Env tilt compared to the relatively planar bilayer. It would be helpful to have an additional panel somewhere that shows the orientation of the antibody (e.g., a representative pose) in the simulations relative to an appropriately curved membrane, Env, the binding conformation of the antibody to Env, and apo Env, given the tilting observed in PMID: 32348769 and theorized in PMC5338832. What additional conformational changes or tilting need to occur between the antibodies and Env to accomplish binding to their respective epitopes?

Thank you for outlining an interesting element to consider in our analysis of a multi-step binding mechanism for MPER antibodies. We added additional figure panels in the supplement to outline the similarities and differences between our simulations and Fabs with the inferred membranes in cryo-EM experiments of full-length HIV Env. The simulated Fabs’ angles are very similar with only minor tilting to match the cryo-EM antibody-membrane geometries.

We added Figure 1-figure supplement 1A & Figure 2-figure supplement 2A, and alter to text to reflect this:

“The primary difference is Env-bound Fabs in cryo-EM adopt slightly more shallow approach angles (~15_°_) relative to the bilayer normal. The simulated bnAbs in isolation prefer orientations slightly more upright, but presenting CDRs at approximately the same depth and orientation. Thus, these bnAbs appear pre-disposed in their membrane surface conformations, needing only a minor tilt to form the membrane-antibody-antigen neutralization complex.”

Env tilt dynamics and membrane curvature of natural virions may reconcile some of these differences. Recent in situ tomography of Full-length Env in pseudo-virions corroborates our approximation of flat bilayers over the short length scales around Env.

The sentence "we next examined the geometries" mentions "potential energy cost, if any, for reorienting...". However, there's no further discussions of geometry or energy cost within this section. Please rephrase, or move this figure to main and increase discussion associated with the various conformational ensembles, their geometry, and their phospholipid association.

As the reviewer highlights, the unbiased simulations and our analysis do not explicitly evaluate energetics. We removed this phrase, and now only allude to the minimal energy barrier between the similar geometric conformations, relative to the tilting & access requirements for antigen binding mechanism.

“The apparent barrier for re-orientation is likely much less energetically constraining than shielding glycans and accessibility of MPER”

".. describing the spectrum of surface-bound conformations" cites the wrong figure.

Thank you for noticing this error; we correct the figure reference to (Figure 2-figure supplement 4).

Please comment on the significance of how global clustering (Fig. S5A-C) was similar for 4E10 and PGZL1, but different for 10E8 (e.g., blue, orange, and yellow clusters for 4E10 and PHZL1 versus cyan, red, and green clusters for 10E8). As the cyan cluster seems to be much closer in Euclidian space to the 4E10/PGZL1 clusters, it might warrant additional analysis. What do these clusters represent in terms of structure/conformation? How do these clusters differ in membrane insertion as in (A)?

We are grateful you identify analysis in the geometric clustering section that may be of interest to other readers. We have added additional supplementary table (Table 2) to detail the CDR loop membrane insertion and global Fab angles which describe each cluster, to demonstrate their similarities and differences. We also better describe how global clustering was similar for 4E10 and PGZL1, but different for 10E8 in the relevant results section

The cyan cluster is not close in structure to 4E10/PGZL1 clusters. We note the original figure panel had an error. The updated Figure 2-supplement 4B shows the correct Euclidian distance hierarchy with an early split between 4e10/pgzl1 and 10e8 clusters.

Figure 3 main text

The start of this section, "We next studied bnAb LN01...", is a good place for a new subheader.

We have added an additional subheader here: Antigen influence on membrane bound conformations and lipid binding sites for LN01

There should be a sentence in the main text defining the replicate setup and production MD run time. Is the apo and complex based on a published structure? How do you embed the MPER? Is the apo structure docked to membrane like in 4E10? The MD setup could also be better delineated within the methods.

The first two paragraphs in this section have been updated to clarify the relevant simulations configuration and Fab membrane docking prediction details.

The procedure was the same for predicting an initial membrane insertion, albeit now we use the LN01-TM complex and the calculation will account for the membrane burial of the the TM domain and MPER fragment. As mentioned, LN01 is predicted as inserted with CDR loops insert similarly with or without the TM-MPER fragment. The geometry differs from PGZL1/4E10 and 10E8, denoted by the text.

Please comment on the oligomerization state of the antigen used in the MD simulation: how does the simulation differ from a crossed MPER as observed in an MPER antibody-bound Env cryo-EM structure (PMID: 32348769), a three-helix bundle (PMC7210310), or single transmembrane helix (PMC6121722)? How does the model MPER monomer embed in the membrane compared to simulations with a trimeric MPER (PMC6035291, PMID: 33882664)-namely, key arginine residues such as R696?

We thank the reviewer for pointing out critical underlying rationale for modeling this TM-MPER-LN01 complex which we have corrected in the revised draft. The range of potential conformations and display of MPER based on TM domain organization could easily be its own paper – we in fact have a manuscript in preparation on the topic.

The updated text expands the rationale for choosing the monomeric uninterrupted helix form of the MPER-TM model antigen (para 1 of LN01 section). The alternative conformations we did not to explore are called out, with references provided by the reviewer.

The discussion qualified that the MPER presentation is likely oversimplified here, noting MPER display in the full-length Env trimer will vary in different conformational states or membrane environments. However, the only cryo-EM structures of full-length ENV with TM domains resolved have this continuous helix MPER-TM conformation – seen both within crossing TM dimers or dissociated TM monomers.

Are there additional analyses that can validate the dynamics of the MPER monomer in the membrane and relative to LN01? Such as key contacts you would expect to maintain over the duration of the MD simulation?

We also increased description of this TM domain’s behavior, dynamics (tilt, orientation, Arg696 snorkeling, and complex w LN01) to provide a clearer picture of the simulation results – which aligns with past MD of the gp41 TM domain as a monomer (para 2 of LN01 section). As well, we noted key LN01-MPER contacts that were maintained.

How does the model MPER modulate membrane properties like lipid density and lipid proximities near LN01?

We checked and didn’t notice differences for the types of lipids (chol, etc) proximal to the MPER-TM or the CDR loops versus the bulk lipid bilayer distributions. Due to the already long & detailed nature of this manuscript, we elect not to include discussion on this topic.

Supplemental figure 1H-I would be better positioned as a figure 3-associated supplemental figure.

We rearranged to follow the eLife format and have paired supplemental panels with their most relevant main figures.

Figure 3F/H reference a "loading site" but this site is defined much later in the text, which was confusing.

Thank you for pointing out this source of confusion, we rearranged our discussion to reflect the order in which we present data in figures.

What evidence suggests that lipids "quickly exchange from the Loading site into the X-ray site by diffusion"? I do not gather this from Figure S1H/I.

We have rearranged the loading side and x-ray site RMSD maps in Figure 3-Figure supplement 1 to better illustrate how a lipid exchanges between these sites.

Figure 4 main text

The authors assert that in the CG simulations, restraints, "[maintain] Fab tertiary and quaternary structure". However, backbone RMSD does not directly assert this claim-an additional analysis of the key interfacial residues between chains, or geometric analysis between the chains, would better support this claim.

Thank you for pointing this point. We rephrased to add that the major sidechain contacts between heavy and light chain persist, in addition to backbone RMSD, to describe how these Fabs maintain the fold stably in CG representation.

In several cases, CG models sample and then dissociate from the membrane. In the text, the authors mention, "course-grained models can distinguishing unfavorable and favorable membrane-bound conformations". Is there a particular orientation that causes/favors membrane association and dissociation? This analysis could look at conformations immediately preceding association and dissociation to give clues as to what orientation(s) favor each state.

Thank you for suggesting this interesting analysis. Clustering analysis of associated states are presented in Figure 5, Figure 5-Figure Supplement 1, and Figure 6, which show all CDR and framework loop directed insertion. This feature is currently described in the main text.

We did not find strong correlation of specific orientations as “pre-dissociation” states or ineffective non-inserting “scanning” events. We revised the key sentence to reflect the major take away – that non-CDR alternative conformations did not insert and most of those having CDRs inserted in a different manner than all-atom simulations also were prone to dissociate:

“Given that non-CDR directed and alternative CDR-embedded orientations readily dissociate, we conclude that course-grained models can distinguish unfavorable and favorable membrane-bound conformations to an extent that provides utility for characterizing antibody-bilayer interaction mechanisms.”

Figure 6 main text

"For 4E10, trajectories initiated from all three geometries..." only two geometries are shown for each antibody. Please include all three on the plot.

The plots include markers for all three geometries for 4E10, highlighted in stars or with letters on the density plots of angles sampled (Figure 6B,C)

"Aligning a full-length IgG... unlikely that two Fabs simultaneously..." Are there theoretical conformations in which two Fabs could simultaneously associate with membrane? If this was physiological or could be designed rationally, could an antibody benefit further from avidity?

Our modeling suggests the theoretical conformations having two Fabs on the membrane are infeasible. It’s even less likely multiple Env antigens could be engaged by one IgG. We have revised the text to express this more clearly.

Figure 7 main text

"An intermediate... showed a modest reduction in affinity..." what affinity does PGZL1 have for this antigen?

The preceding sentence for this information: “Mature PGZL1 has relatively high affinity to the MPER epitope peptide (Kd = 10 nM) and demonstrates great breadth and potency, neutralizing 84% of a 130 strain panel “

Figures

Figure 1

It would be helpful to have an additional panel at the top of this figure further zoomed out showing the orientation of the antibody (e.g., a representative pose) in the simulations relative to an appropriately curved membrane, Env, the binding conformation of the antibody to Env, and apo Env, given the tilting observed in PMID: 32348769 and theorized in PMC5338832. What additional conformational changes or tilting need to occur between the antibodies and Env to accomplish binding to their respective epitopes?

Thank you for the suggestion to include this analysis. We have added to the text reflecting this information, as well as making new supplemental panels for 4E10 and 10E8 that we compare simulated 4E10 and 10E8 Fab conformations to cryoEM density maps with Fabs bound to full-length HIV Env. Figure 1-figure supplement 1A & Figure 2-figure supplement 2A

In Figure 1, space permitting, it would be helpful to annotate the distances between the phosphates and side chains (similarly, for Figure S1A).

To avoid the overloading the Main figure panels with text, those relevant distances are listed in the methods sections. Those distances are used to define the “bound” lipid phosphate state. Generally, we note the interactions are within hydrogen bonding distance.

Annotating "Replicate 1" and "Replicate 2" on the left side of Figure 1C/D would make this figure immediately intuitive.

We have added these labels.

Figure caption 1C: Please clarify the threshold/definition of a contact used to binarize "bound" versus "unbound" (for example, "mean distance cutoff of 2A between the phosphate oxygen and the COM of CDR-H1") [on further reading of the methods section, this criterion is quite involved and might benefit from: a sentence that includes "see methods"]. Additionally, C could use a sentence explaining the bar such as in E, "Phosphate binding is mapped to above each MD trajectory" Please define FR-H3 in the figure caption for E/F.

We have added these details to the figure caption.

Because Figure 1 is aggregated simulation time, it would be helpful to also represent the data as individual replicates or incorporate this information to calculate standard deviations/statistics (e.g., 1 microsecond max using the replicates to compute a standard deviation).

We believe the current quantification & display of data via sharing all trajectories is sufficient to convey the major point for how often each CDR-phosholipid binding site it occupied. Further tracking and statistics of inter-atomic distances will likely be too tedious & add minimal value. There is some dynamics of the phosphate oxygens between the polar within the CDR site but our “bound” state definitions sufficiently describe the key participating interactions are made.

Figure 2

For A, it would be helpful to annotate the yellow and blue mesh on the figure itself.

We have defined the orange phosphate and blue choline densities.

Also, where are R29 and Y32 relative to this site? In the X-ray panels, Y38 is not shown, and the box delineating the zoom-in is almost imperceptible.

Thank you for this suggestion to include those amino acids which are referenced in the text as critical sites where mutation impacts function. To clarify, Y32 is the pdb numbering for residue Y38 in IMGT numbering. We have added a panel to Figure 2-Figure Supplement 1 having a cartoon graphic of 10E8 loop groove with sidechains & annotating R29 and Y38, staying consistent with out use of IMGT numbering in the manuscript.

Figure 3

It might read clearer to have "LN01+MPER-TM" and "LN01-Apo" in the middle of A/B and C/D, respectively, and a dotted line delineating the left and right side of the figure panels.

We have added these details to the figure for clarity for readers.

It would be helpful to show some critical interactions that are discussed in the text, such as the salt bridge with K31, by labeling these on the figure (e.g., in E-H).

We drafted figure panels with dashed lines to indicate those key interactions. However, they became almost imperceptible and overloaded with annotations that distracted from the overall details. For K31, the interaction occurs in LN01 crystal structures readers can refer to.

Why are axes cut off for J?

We corrected this.

Please re-define K/L plots as in Figure 1, and explain abbreviations.

We updated the figure caption to reflect these changes.

Figure 4

The caption for panel A states that the Fab begins in solvent 1-2 nm above the bilayer, but the main text states 0.5-2 nm.

We have reconciled this difference and listed the correct distances: 0.5-2nm.

Please label the y-axis as "Replicate" for relevant figure panels so that they are more immediately interpretable.

This label has been added.

A legend with "membrane-associated" and "non-associated" within the figure would be helpful. Additionally, the average percent membrane associated, with a standard deviation, should be shown (Similar to 1C, albeit with the statistics).

This legend has been added. We also added the additional statistical metrics requested to strengthen our analysis.

The text references "10, 14, and 12 extended insertion events" for the three antibody-based simulations. How do you define "extended insertion events"? Would breaking this into average insertion time and standard deviation better highlight the association differences between MPER antibodies and controls, in addition to the variability due to difference random initialization?

We thank the reviewer for the insightful suggestion on how to better organize quantitative analysis to support the method. Supplemental Table 3 includes these numbers.

Figure 5

The analysis in Fig. S6C could be included here as a main figure.

The drafted revised figure adding S6C to Figure 5 made for too much information. Likewise, putting this panel S6C separated it from the parent clustering data of S6B, so we decided to keep these figures separated. The S6 figure is now Figure 5-figure supplement 1.

Figure 6

Please annotate membrane insertion on E as %.

These are phosphate binding RMSD/occupancy vs time. The panels are now too small to annotate by %. The qualitative presentation is sufficient at this stage. The quantitative % are listed in-line within text when relevant to support assertions made.

Please use the figure caption to explain why certain clusters (e.g., 10E8 cluster A, artifact, Fig. S6E) are not included in panel E.

We have added this information in the figure caption.

Figure 7

Please show all points on the box and whisker plots (panels E and F), and perform appropriate statistical tests to see if means are significantly different (these are mentioned in the text, but should be annotated on the graph and mentioned within the figure caption).

We have changed these plots to show all data points along with relevant statistical comparisons. The figure captions describe unpaired t-test statistical tests used.

Figure S1

G, H, and I do not belong here-they should be moved to accompany their relevant text section, which associates with Figure 3. It would be helpful to associate this with Figure 3 in the eLife format, "Figure 3-Supplemental Figure 1" or its equivalent.

It's very difficult to distinguish the green and blue circles on panel G.

We darkened the shading and added outline for better visualization

Subfigure I is missing a caption, could be included with H: "(H,I) Additional replicates for LN01+TM (H) and LN01 (I)".

We corrected this as suggested.

Why is H only 3 simulations and not 4? Does it not have a lipid in the x-ray site? Also, the caption states "(top, green)" and "(bottom, cyan)", but the green vs. cyan figures are organized on the left and right. Additional labels within the figure would help make this more intuitive.

If the point of H and I is to illustrate that POPC exchanges between the X-ray and loading sites, this is unclear from the figure. Consider clarifying these figures.

Thank you for describing the confusion in this figure, we have added labels to clarify.

Figure S2 (panels split between revised Figure 4 associated figure supplements)

The LN01 figures should likely follow later so that they can associate with Figure 3, despite being a similar analysis.

We corrected supplements to eLife format so supplements are associated with relevant main figures.

Figure S3 (panels split between revised Figure 1 & 2 associated figure supplements)

As hydrophobicity is discussed as a driving factor for residue insertion, it would be helpful to have a rolling hydrophobicity chart underneath each plot to make this claim obvious.

We prefer the current format, due to the worry of having too much information in these already data-rich panels. As well, residues are not apolar but are deeply inserted.

Figure S4 (panels split between revised Figure 1 & 2 associated figure supplements)

It would be helpful to label the relevant loops on these figures.

We have labeled loops for clarity.

Do any of these loops have minor contacts with Env in the structure?

The 4E10 and PGZL1 CDRH-1 loop does not directly contact bound MPER peptides bound in crystal structures.

FRL-3 and CDR-H1 in 10E8 do not contact the MPER peptide antigen component based on x-ray crystal structures.

Do motif contacts with lipid involve minor contacts with additional loops other than those displayed in this figure?

The phosphate-loop interactions in motifs used as query bait here are mediated solely by the backbone and side chain interactions of the loops displayed. We visually inspected most matches and did not see any “consensus” additional peripheral interactions common across each potential instance in the unrelated proteins. The supplied Supplemental Table 2 contains the information if a reader wanted to conduct a detailed search.

Why is there such a difference between the loop conformation adopted in the X-ray structure and that in the MD simulation, and why does this lead to the large observed differences in ligand-binding structure matches?

We thank the reviewer for carefully noting our error in labeling of CDR loop and framework region input queries. We revised the labeling to clarify the issue.

The is minimal structural difference between the loops in x-ray and MD.

Figure S5 (Figure 2-Figure supplement 4)

This figure is not colorblind friendly-it would be helpful to change to such a pallet as the data are interesting, but uninterpretable to some.

We have left this figure the same.

"Susbstates" - "Substates"

Corrected, thank you.

Panel B is uninterpretable-please break the axis so that the Euclidian distances can be represented accurately but the histograms can be interpreted.

We have adjusted axis for this plot to better illustrate the cluster thresholds.

The clusters in D-H should be analyzed in greater depth. What is the structural relevance of these clusters other than differences in phospholipid occupancy in (I)? Snapshots of representative poses for each cluster could help clarify these differences.

We have adjusted the text to describe the geometric differences in each of those clusters that result in the different exceptionally lower propensities for forming the key phospholipid interaction.

The figure caption should make it clear that 3 μS of aggregate simulation time is being used here instead of 4 μS to start with unique tilt initializations. E.g., "unique starting membrane-bound conformations (0 degrees, -15 degrees, 15 degrees initialization relative to the docked pose)". Further, why was the particular 0-degree replicate chosen while the other was thrown out? Or was this information averaged? Why is the full 4 μS then used for D-I?

We thank the reviewer for noting these details. We didn’t want to bias the differential between 10E8 and 4E10/PGZL1 by including the replicate simulations. The analysis was mainly intended to achieve more coarse resolution distinction between 10E8 and the similar PGZL1/4E10.

In the subsequent clustering of individual bnAb simulation groups, the replicate 0 degree simulations had sufficiently different geometric sampling and unique lipid binding behavior that we though it should be used (4 us total) to achieve finer conformational resolution for each bnAb.

Figure S6 (now Figure 5-Figure Supplement 1)

Please label the CDRs in C and provide a color key like in other figures. Also, please label the y-axes. This figure could move to main below 5B with the clusters "A,B,C" labeled on 5B.

We have added the axes labels and color key legend. We retained a minimal CDR loop labeling scheme for the more throughput interaction profiles here where colored sections in the residue axes denote CDR loop regions.

Figure S7 (Figure 7 Figure Supplement 1)

Panels A and B would likely read better if swapped.

We have swapped these panels for a better flow.

For panel C, please display mean and standard deviation, and compare these values with an appropriate statistical test.

This is already displayed in main figure, we have removed it from supplement.

For E and F, please clarify from which trajectory(s) you are extracting this conformation from. Are these the global mean/representative poses? How do they compare to other geometrically distinct clusters?

The requested information was added to supplemental figure caption. These are frames from 2 distinct time points selected phosphate bound frames from 0-degree tilt replicates for both 4E10 and 10E8, representing at least 2 distinct macroscopic substates differing in global light chain and heavy chain orientation towards the membrane.

Table S2 (now Supplementary Table 3)

Please add details for the 13h11 simulation.

Additionally, please add average contact time and their standard deviation to the table, rather than just the aggregated total time. This will highlight the variability associated with the random initializations of each simulation.

We have added the details for 13h11 and the requested analysis (average aggregated time +/- standard deviation and average time per association event +- standard deviation) to supplement our summary statistics for this method.

Reviewer #2 (Recommendations For The Authors):

(1) The structure of the manuscript should be improved. For example, almost half of the introduction (three paragraphs) summarize the results. I found it hard to navigate all the data and specific interactions described in the result section. Furthermore, the claims at the end of several sections seem unsupported. Especially for the generalization of the approach. This should be moved to the discussion section. The discussion is pretty general and does not provide much context to the results presented in this study.

We have significantly reorganized the results section to improve the flow of the manuscript and accessibility for readers, especially the first sections of all-atom simulations. We also removed claims not directly supported by data from our results, and expanded on some of these concepts in the discussion to make some more novel context to the result.

(2) The author should cite more rigorously previous work and refrain from using the term "develop" to describe the simple use of a well established method. E.g. Several studies have investigated membrane protein interactions e.g. [1], membrane protein-bilayer self-assembly [2], steered molecular dynamics [3], etc.

Thank you for identifying relevant work for the simulations that set precedent for our novel application to antibody-membrane interactions. We have removed language about development of simulation methods from the text and now better reference the precedent simulation methods used here.

(3) Have the authors considered estimating the PMF by combining the steered MD simulation through the application of Jarzynski's equality?

We performed from preliminary PMFs for Fab-membrane binding, but saw it was taking upward of 40 us to reach convergence. Steered simulations focus on a key lipid may be easier.

Although PMFs are beyond the scope of this work, we added proposals & allusion to their utility as the next steps for more rigorous quantification of fab-membrane interactions.

Minor

(4) The term "integrative modeling" is usually used for computational pipelines which incorporate experimental data. Multiscale modeling would be more appropriate for this study.

We altered descriptions throughout the manuscript to reflect this comment.

(5) Units to report the force in the steered molecular dynamics are incorrect. They should be 98.

We changed axes and results to correctly report this unit.

(6) Labels for axes of several graphs are not missing.

We added labels to all axes of graphs, except for a few where stacked labels can be easily interpreted to save space and reduce complexity in figures.

(7) Figure 3 K & L is this really < 1% of total? The term "total" should also be clarified.

Thank you for pointing this out, we changed the % labels to be correct with axes from 0-100%. We clarified total in the figure caption.

(8) The font size in figures should be uniformized.

This suggestion has been applied

(9) Time needed for steered MD should be reported in CPUh and not hours (page 17).

We removed comments on explicit time measurements for our simulations.

(10) Version of Martini force field is missing in methods section

We used Martini 2.6 and added this to the methods.

References

(1) Prunotto, Alessio, et al. "Molecular bases of the membrane association mechanism potentiating antibiotic resistance by New Delhi metallo-β-lactamase 1." ACS infectious diseases 6.10 (2020): 2719-2731.

(2) Scott, Kathryn A., et al. "Coarse-grained MD simulations of membrane protein-bilayer self-assembly." Structure 16.4 (2008): 621-630.

(3) Izrailev, S., et al. "Computational molecular dynamics: challenges, methods, ideas. Chapter 1. Steered molecular dynamics." (1997).

Version 2

Abstract

Summary

A potent class of HIV-1 broadly neutralizing antibodies (bnAbs) targets the envelope glycoprotein’s membrane proximal exposed region (MPER) through a proposed mechanism where hypervariable loops embed into lipid bilayers and engage headgroup moieties alongside the epitope. We address the feasibility and determinant molecular features of this mechanism using multi-scale modeling. All-atom simulations of 4E10, PGZL1, 10E8 and LN01 docked onto HIV-like membranes consistently form phospholipid complexes at key complementarity-determining region loop sites, solidifying that stable and specific lipid interactions anchor bnAbs to membrane surfaces. Ancillary protein-lipid contacts reveal surprising contributions from antibody framework regions. Coarse-grained simulations effectively capture antibodies embedding into membranes. Simulations estimating protein-membrane interaction strength for PGZL1 variants along an inferred maturation pathway show bilayer affinity is evolved and correlates with neutralization potency. The modeling demonstrated here uncovers insights into lipid participation in antibodies’ recognition of membrane proteins and highlights antibody features to prioritize in vaccine design.

Introduction

Antibodies can target integral membrane proteins very near to the lipid bilayer surface, including at epitopes partially embedded within the headgroup region. During antigen engagement, antibodies complementarity-determining regions (CDR) may need to navigate lipid molecules to gain access to a concealed or dynamically buried site. Evolving tolerance or even affinity to lipid bilayers could be beneficial in recognition of membrane-proximal epitopes, potentially forming simultaneous cooperative interactions with antigen and membrane yielding additional avidity and specificity. Conversely, propensity to bind lipids or cell membranes poses a significant auto-immunity risk. B-cells producing antibodies targeting host membranes are downregulated in healthy organisms1–4. Nonetheless, rare antibodies with lipid affinity can emerge, particularly in cases of chronic inflammation and infection as in HIV5–8. Maturation pathways of these rare events remain unclear, but must strike a careful balance of polyreactivity to avoid or overcome autoreactivity9. A better understanding of how antibodies develop membrane affinity and target membrane-proximal epitopes would be impactful for antibody therapeutics, auto-immunity, and vaccine development10–14.

We sought to address this phenomenon for broadly neutralizing antibodies (bnAbs) 4E10, PGZL1, 10E8, and LN01 which stem from unique lineages and all target the semi-concealed membrane-proximal epitope region (MPER)15–18 of the HIV-1 envelope glycoprotein (Env). Interestingly, these bnAbs show varying degrees of affinity for lipid components; some associate with lipid bilayers or cultured cells even in the absence of antigen5–819–25. This membrane interaction behavior appears to correlate with neutralization potency and is attributed to shared CDR loop features including a long hydrophobic CDR-H37,20,23,24,26–28. The correlation between membrane association and neutralization has been established, given that mutations or chemical conjugations to solvent-exposed antibody loop residues which enhance association to phospholipid vesicles consequently boost efficacy of pseudovirus neutralization29. Conversely, mutations reducing hydrophobicity drastically reduce neutralization activity while also weakening association to lipid bilayers, despite minimal impacts to antigen affinity (e.g. 4E10 CDR-H3 H100-H102 Trp-Trp motif; PGZL1 clone H4K3’s CDR-H1)7,23,30,31. Thus, phospholipid membrane interactions evolved by MPER-targeting bnAbs appear critical for their mechanism of immune protection against HIV. Yet, the breadth of molecular features responsible for association with membranes and mode of lipid-embedded epitope engagement are not clear.

Structural characterization of MPER-targeting bnAbs with full-length Env trimer or gp41 fragments also indicate that the surrounding lipid bilayer plays a role in antibody access and epitope recognition15,32,33. Cryo-electron microscopy (cryo-EM) of Env trimers bound to bnAbs PGZL1, 4E10, and 10E8 within in different model membranes suggest their CDR loops form intimate contacts with surrounding lipids while engaging MPER32,33. Antigen binding fragments (Fab) crystal structures of those bnAbs as well as LN01 soaked with short-chain phospholipids all revealed ordered headgroup moieties complexed within the CDR loops in the presence and absence of antigen5,8,22,34,34 suggesting the antibodies encode specific lipid interactions5,8,22,34,34. Thus, the molecular features mediating membrane affinity for these bnAbs appear critical to their maturation and mechanism of immune protection against HIV in vivo.

These data support a 2-step bnAb neutralization mechanism proposed previously, wherein a population of bnAbs in vivo may first associate with membranes via embedding their CDRs, then laterally diffuse across the bilayer surface to subsequently engage Env at MPER6,21,24,26,27,32,34–36. The reverse order, MPER binding followed by CDR membrane insertion, is also often posited. Here, we developed multi-scale molecular dynamics (MD) simulation approaches suited to investigate these mechanisms and the unique maturation landscape these rare antibodies must navigate to avoid auto-immune consequences. For these MPER bnAbs, we focused on the in-silico capacity to characterize the molecular features mediating phospholipid affinity at atomic detail, not afforded by previous structural approaches. We find that unbiased all-atom MD simulations accurately and reproducibly predict ab initio binding of phospholipids stably bound at specific CDR binding sites previously identified in co-crystal structures21,22,26,34. Integrating coarse-grain (CG) simulations, we capture the full process of antibodies scanning and embedding into lipid bilayers to biologically relevant conformations. The simulations illuminate key molecular features tuning bnAbs’ lipid interaction and preferred membrane-bound geometry, advancing on previous work35, including surprising contributions from framework regions. Biased simulations provide rough estimates of membrane interaction strength, applied to demonstrate that in silico bilayer affinity correlates with neutralization efficacy of experimentally characterized PGZL1 variants along a pseudo-maturation trajectory18.

This modeling platform provides an improved framework for navigating the proposed 2-step bnAb neutralization mechanism both for retrospective and proactive evaluation of antibody repertoires, which should have utility in vaccine design and therapeutic antibody engineering. Likewise, our results imply more generalized implications for the in vivo selection of antibodies targeting membrane proteins, beyond HIV, at partially concealed juxtamembrane regions – highly desirable epitope regions given their typically high protein sequence conservation. Reliable atomic detail views of antibody conformations at the membrane surface and loop phospholipid binding should be powerful for better understanding membrane protein epitope engagement, pathogen neutralization mechanism, and the checkpoints regulating self-sensing antibodies.

Results

Atomic simulations accurately model MPER bnAb 4E10 and PGZL1 phospholipid interactions at HIV-like membrane surfaces

To assess the stability, organization, and spectrum of phospholipid interactions inferred from previous lipid-bound crystallography and cryo-EM experiments, we began by performing unbiased all-atom simulations of 4E10, PGZL1, and 10E8 peripherally embedded to the surface of model lipid bilayers. Initial membrane-bound Fab conformations were computed using PPM2.03735, globally optimizing protein insertion based on per-residue hydrophobicity and solvation, each resulting in reasonable predictions with CDR-H3 embedded. Four 1 μs pseudo-replicate simulations were initiated per bnAb, with two replicates tilted by ±15 degrees to modestly vary the starting membrane-interacting pose, utilizing an explicit simplified HIV-like model anionic bilayer: 25% cholesterol, 5% 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphate (POPA), 70% 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC)6.

4E10 and PGZL1 are homologous (85% sequence identity, sharing IgG VH1-69, VH4-34 and VH4-59), and have very similar affinities, breadth, and potency (especially comparing PGZL1.H4K3). 4E10 is notably more poly-reactive: binding to cultured cells, numerous isolated phospholipids5, and vesicles of most compositions even in the absence of antigen – whereas 10E8, LN01, and PGZL1 do not. PGZL1 and 4E10 crystal structures both bear putative lipid electron density at several surface loops, notably recurring within a CDR-H1 site – a region when mutation reduced the neutralization potency 10-fold for PGZL1-H4K322,34,38. Given their similarities, we first compared the lipid bilayer interaction mechanisms in 4E10 and PGZL1 simulations. Upon time-averaging phospholipid density from 4 independent trajectories, both bnAbs showed strong hotspots for a lipid phosphate bound within the CDR-H1 loops, with minimal phospholipid or cholesterol ordering around the proteins elsewhere. The simulated lipid phosphates bound within CDR-H1 have exceptional overlap with electron densities and atomic details of modelled headgroups from respective lipid-soaked co-crystal structures of 4E10 and PGZL122,26,34, with phosphate mean positions RMSDs of 0.6 and 0.7 Å, respectively, relative to the X-ray structures (Figure 1A-B; Figure Supplement 1A-D). Thus, our simulations reproduce ab initio with atomic accuracy the evidence of 4E10 and PGZL1 bnAbs’ mechanism of membrane association. These results validate the CDR-H1 antibody-lipid interactions proposed from previous co-crystal structures are not only feasible in the context of full biologically realistic HIV-like bilayers, but form very readily. The CDR-H1 lipid phosphate complexes appear to be specific interactions that anchor and orient CDR loops at the bilayer surface for antigen engagement, complementing protein-lipid interactions which are less geometrically specific, i.e., lipid tail solvation of apolar groups in the CDR-H loops such as CDR-H3 tryptophans.

4E10 and PGZL1 CDR lipid phosphate binding sites and membrane interaction profiles in all-atom fluid HIV-like bilayer MD simulations

(A) Representative frame from an MD simulation of the phospholipid complex at 4E10 Fab CDR-H1 (top left). Top right, time-averaged lipid phosphate density (orange mesh) relative to antibody CDR loops embedded in the bilayer (beige) from 4 µs total time (top right). Center right, CDR-H1 loop side chain and backbone atoms of de novo predicted phosphate interaction (middle right). Bottom, putative lipid phosphate binding site observed in an X-ray structure for 4E10 (PDB: 4XC1), comparing the CDR-H1 loop interactions to MD (bottom right).

(B) PGZL1 lipid phosphate interaction at the CDR-H1 loop from MD simulation versus X-ray crystallography (PDB: 6O3J), demarked as in (A).

(C) RMSD of the interacting lipid phosphate versus the experimental CDR-phosphate position (X-ray site), classified as bound (green) and unbound (grey) by loop-phosphate contacts (see methods) in 2, 4E10 representative replicate trajectories. Black line, ten-frame RMSD running average; standard deviation, grey shading.

(D) RMSD of lipid phosphate binding to PGZL1 CDR-H1 in MD simulations versus X-ray site. Phosphate binding is mapped above each MD trajectory as in (C).

(E) Per-residue interaction profiles for Fab simulations of 4E10 detailing the time spent for each residue in in phosphate layer (orange), glycerol layer (red), or hydrocarbon layer (blue) across aggregate 4 µs from 4 simulations. CDR loops are mapped in solid color blocks below each profile. Fab domain regions making significant contact are labeled, including CDRs and heavy chain framework region 3 (FR-H3).

(F) Per-residue interaction profiles for antibody Fab simulations for PGZL1, colored as in (E).

We next characterized the stability and dynamics of the spontaneously forming antibody-phospholipid complexes which recurred across all trajectories. For both 4E10 and PGZL1, 3 of 4 replicates showed very rapid binding of a headgroup phosphate ab initio in the first 20 nanoseconds (ns) from lipids diffusing in from >= 8 Å away) (Figure 1C-D). In some cases, phospholipids bound as early as the protein-restrained equilibration (15 ns) where lipids can freely diffuse (Figure 1—figure supplement 1D-E). For both 4E10 and PGZL1, one of each replicate simulation had delayed formation of the stable CDR-H1-mediated phospholipid complexes (within 200 ns) requiring minor Fab reorientation on the membrane surface given these replicates’ initialization from an artificially tilted orientation (Supp. Video 1). Upon formation, most phospholipid complexes were highly stable, typically persisting for hundreds of nanoseconds with low mobility of the phosphate headgroup relative to CDR-H1 (<1 Å RMSF) and high overall occupancy (both >80% across 4 microseconds total). An extensive polar network coordinates the lipid phosphate oxygens within the CDR-H1 loop: hydrogen bonds donated from backbone amides of Phe29 and Ser30 and from sidechain hydroxyls of Ser29/Thr29 and Ser30 (Figure 1C-D)22,34.

Notably, CDR-H loops in contact with the membrane maintained internally rigid backbone conformations during simulation with CDR-H1 having < 1 Å RMSF (Figure 1–figure supplement 1C-D). One PGZL1 simulation captured a rare dynamic lipid exchange event (Figure 1—figure supplement 1E, Supp. Video 2). A POPC molecule, bound for ∼500 ns, dissociated from the CDR-H1 and was promptly replaced by a second nearby lipid, POPA. These observations, including the kinetic stability and high occupancy, suggest that 4E10 and PGZL1 family antibodies likely are predominantly bound to an anchoring phospholipid when intimately contacting the membrane surface, including during MPER engagement and in the neutralized complex. When CDR-H3 is inserted, CDR-H1 is presented as a rigid pre-formed binding site conveniently at the membrane’s headgroup layer for rapid formation of complexes, with those phospholipids likely in a steady-state equilibrium freely exchanging (nano-microsecond timescale) yet heavily favoring the bound state.

These results of reliably recovering experimentally determined CDR-phospholipid complexes bolsters confidence that antibody behaviors within the simulations are biologically relevant, motivating our deeper analysis of features driving the bnAbs’ protein-lipid interactions and global membrane surface conformations. Characterizing the cumulative per-residue interaction profiles of 4E10 and PGZL1 with bilayer lipids categorized into headgroup, glycerol, or hydrocarbon layers illustrates the deep immersion of heavy-chain loops, with the light-chain making some peripheral polar headgroup contacts (namely CDRs L1 and L3) (Figure 1E-F). The patterns of lipid interactions compare well with previous 4E10 simulations35,39, yet are more expansive here – possibly due to the sustained phospholipid-CDR-H1 complex and greater simulation length.

Aligning interaction profiles with corresponding primary sequences contextualizes the amino acid molecular features responsible and their origin in antibody maturation. One novel observation consistent between our 4E10 and PGZL1 simulations was that heavy-chain framework region 3 (FR-H3) extensively intercalates into the bilayer surface. This finding demonstrates that predominantly germline encoded antibody surface features can provide significant contributions to the membrane binding mechanism and surface orientation of antibodies – a key insight for invoking and honing antibody-lipid interactions in vaccine design. For the light-chain loops and FR-H3 of PGZL1 and 4E10, charged residues mediate their shallower headgroup surface interactions (Figure 1-figure supplement 2A-B), whereas the CDR-H loops bare neighboring stretches of hydrophobic and lipophilic polar residues facilitating their deeper embedding. Compared to PGZL1, 4E10’s CDR-H loops are further buried, more extensively contacting lipid aliphatic tails – which may in part explain 4E10’s higher tendency for poly-specificity. The comprehensive protein-lipid atomic-detail maps further detail that the mechanisms for membrane interaction and antibody maturation driven are largely by the solvation for apolar surface loops – the extent of hydrophobic burial – complemented by compatible polar residue electrostatics. The majority of bnAb protein-lipid interactions arise from surface residues’ macroscopic chemical properties, contrasting the much more sequence and conformationally specific phospholipid binding sites constructed at CDR-H1. Thus, the multivalent combination of surface features combined with highly specific lipid anchoring sites may be a convergent theme for maturation and mechanism for MPER-targeting bnAbs, including the previously undervalued contribution from germline-encoded framework regions to lipid association.

Phospholipid complex formed in the CDR Light Chain groove of 10E8 bound to membrane surfaces

We similarly investigated bnAb 10E8, which differs in its genetic origins and expected light-chain-mediated membrane binding mode34,36,40,41. In all 4 replicate simulations, we observed a POPC complexed at a groove between CDR-L1 and FR-L3 (Figure 2A-B) which had had significant non-protein electron density in previous 10E8 X-ray structures with and without lipids, which had been modelled as headgroup phosphoglycerol anions, glycerol, or free dihydrogen phosphate26. In our simulations, the stably bound POPC phosphate was slightly offset (2.7 Å) from the proposed CDR-L1 crystallographic site (Figure 2-figure supplement 1B), and instead localized to FR-L3 stabilized by a hydrogen bonding network from Ser83 hydroxyl, Gly84 backbone amide, and Asn85 C-alpha proton (Figure 2A). Simultaneously, the POPC choline moiety occupied the precise crystallographic CDR-L1 site, wherein the choline cation was coordinated by four pre-organized unpaired backbone carbonyls stemming from the specific CDR-L1 short loop helix conformation (Figure 2A). We find these bivalent complexes engaging both moieties of the lipid phosphocholine (PC) zwitterion at this CDR-L1 FR-L3 groove with >70% occupancy overall across replicate simulations with high stability: 0.5 Å RMSF for CDR-L1–choline site and 0.5 Å RMSF for FR-L3–phosphate site (Figure 2-figure supplement 1C-D). These 10E8 loops were conformationally rigid (RMSF < 1.0 Å, Figure 2-figure supplement 1E), presenting a pre-formed binding site analogous to CDR-H1 in 4E10 and PGZL1. Previous experiments highlight the importance of CDR-L1 phospholipid binding site features in 10E8’s neutralization mechanism, given that the double mutant of R29 and Y38 to alanine exhibited vastly decreased neutralization (>500 fold) across HIV strains with negligible impact to MPER affinity26. The time-averaged per-residue lipid interaction profile demonstrates how 10E8 uses both heavy (CDR-H3) and light chain loops (CDRs L1, L2, L3; FR-L3) to orient the antibody geometry on the membrane surface to position 10E8’s phospholipid binding groove site at the headgroup-rich bilayer region (Figure 2-figure supplement 1F). These simulations supplement experimental findings with novel atomic details to establish that genetically distinct MPER antibodies share a consistent mechanism of utilizing CDR loops and framework to mediate bilayer interactions, with 10E8 utilizing FR-L3 for phosphate specificity in a stable lipid complex to anchor this bnAb at the membrane surface.

10E8 bivalent lipid headgroup interaction and bilayer insertion predicted in MD simulation closely matches experimental lipid binding sites

(A) Top left, representative frame from MD simulation of lipid interacting with 10E8 Fab. Top right, the MD lipid binding site includes a bivalent choline and phosphate lipid headgroup complex (represented as blue and orange mesh time-averaged positional density in simulations) within a protein surface groove composed of CDR-L1 and FR-L3 respectively. Bottom, positions of phosphates or glycerols modeled at the CDR-L1 and FR-L3 groove site within 10E8 Fab X-ray structures (PDB: 5T85, 5T6L)

(B) RMSD of lipid choline position in the MD simulations versus expected CDR-L1 lipid binding site from 10E8 X-ray structures. “Bound” state assigned relative to choline position and phosphate FR-L3 interactions observed in MD.

(C) Per-residue interaction profiles for antibody Fab simulations for 10E8 with Fab domain regions making significant contact labeled, including CDRs and light chain framework region 3 (FR-L3).

Phospholipid binding modes in complex high cholesterol lipid bilayers

We next evaluated whether such bnAb conformations and ab initio phospholipid binding are robust to distinct lipid compositions in more complex high cholesterol bilayers matching lipidomics studies for HIV particles6,42. New simulations were initialized for similar 4E10, PGZL1, and 10E8 Fab conformations in membranes having a 9.5:19:8.5:18:45 ratio of POPC, 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine (POPE), 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS), palmitoyl sphingomyelin (PSM), Cholesterol. In a 500 ns trajectory for each, stable phospholipid complexes rapidly formed within the respective CDR loop sides with atomic accuracy to the expected position – now having more diversity in lipid headgroup chemistries (Figure 2–figure supplement 1A-C). Although 4E10 bound only POPC at its CDR-H1 site, in PGZL1 simulations CDR-H1 coordinated a PSM molecule (PC headgroup) in precisely the same binding mode as observed for POPC despite PSM’s substitution of two lipid tails off the glycerol backbone. Likewise, 10E8 bound a POPE at its light chain groove site, with the PE (phosphate and cation) forming analogous interactions (<1 Å RMSD to POPC), including coordination of the cationic ethanolamine by the ring of backbone carbonyls capping CDR-L1’s loop helix. Thus, these bnAb phospholipid binding sites can accommodate a broader scope of lipid tail and headgroup chemistries than we have investigated here and can readily form stable phospholipid anchoring complexes in bilayers of HIV-like cholesterol-rich composition. One shared conformational difference observed for these bnAbs the higher cholesterol bilayers was slightly more extensive and broader interaction profiles as well as modestly deeper embedding of the relevant CDR and framework surfaces loops (Figure 2– figure supplement 1D-F). These results bolster the feasibility for using all-atom MD as an in silico platform to explore differential phospholipid affinity at these sites (i.e., specificity studies) and influence on antibody preferred conformation as membrane composition and lipid chemistry are systematically varied. More thorough simulation techniques such as potential of mean force are needed to make those predictions and could better guide experiments concerning how to incorporate the antibody-lipid interaction in neutralization and maturation of bnAbs via vaccine.

Structural bioinformatics contextualizes bnAb interactions at the membrane

We sought to interrogate the biological relevance and in silico predictability of these observed CDR-lipid polar interactions using an orthogonal bioinformatics approach. We mined protein-ligand interactions to assess whether phosphate complexes of similar geometries have been observed in Nature, positing that the occurrence of analogous phosphate interactions outside the context of MPER bnAbs, i.e., within critical structural or functional regions of other protein families, can differentiate common proteome-wide functional motifs from simulation artifacts. Structural searches querying the lipid-binding loop backbone conformation of each Fab (as previously described16,22,26) identified between 105 to 2·106 geometrically similar sub-segments within natural proteins (<2 Å RMSD)43, reflecting they are relatively prevalent (not rare) in the protein universe, comparing well with frequency of other surface loops of similar length in antibodies (Supplementary Table 4). The structurally similar loops found in distinct proteins were subsequently mined for nearby phosphate/phosphoryl and sulfate/sulfo ligands (Figure 2–figure supplement 3; Supplementary Table 1). When searching for loops similar to the CDR-H1 of 4E10 and PGZL1, only 4 cases of phosphate-type ligand binding motifs were identified. Thus, this CDR-H1 site is a realistic but rare protein-phospho-ligand structural motif (Figure 2–figure supplement 3A-B).

For 10E8’s CDR-L1 site, only 1 phosphate-type ligand was observed: a solvent-exposed surface crystallographic ion. By contrast, 10E8’s FR-L3 beta-turn site was a hot spot for phosphate-type ligands, with 23 natural protein cases (Figure 2–figure supplement 3C). Thus, while both CDR-L1 and FR-L3 10E8 loops sites can bind phosphate ligands, the much greater natural prevalence of phosphates ligated at FR-L3 supports that the newly proposed 10E8 POPC complex observed in our simulations is biologically realistic, rather than a simulation artifact. Barring experimental validation, existence of this FR-L3 site establishes precedent for how important framework regions can be for mediating protein-membrane interactions, including hosting stable and specific phospholipid complexes. This bioinformatics analysis illuminates the sophisticated molecular recognition mechanism in which 4E10, PGZL1 and 10E8 Fab domains incorporate recurring structural features from Nature to coordinate specific phospholipid binding interactions.

Global geometries of membrane-bound bnAb and implications for tilted Env engagement

Given the emphasis on antibody tilt relative to the bilayer and antigen in discerning neutralization mechanisms32, we compared the geometries of membrane-bound conformational ensembles for our 4E10, PGZL1, and 10E8 simulations to those of Fab-bound Env trimer cryo-EM structures in membrane mimics32. Simulated antigen-free Fabs favor similar orientations on lipid bilayer surfaces as in bnAb-bound tilted Env experimental structures, likely facilitated by the anchoring phospholipid complexes (Figure 1–figure supplement 1A; Figure 2–figure supplement 1A). The primary difference is Env-bound Fabs in cryo-EM adopt slightly more shallow approach angles (∼15°) relative to the bilayer normal. The simulated bnAbs in isolation prefer slightly more upright orientations, but present CDRs at approximately the same depth and orientation. Thus, these bnAbs appear pre-disposed in their membrane surface conformations, needing only a minor tilt to form the membrane-antibody-antigen neutralization complex. The apparent barrier for re-orientation is likely much less energetically constraining than shielding glycans and accessibility of MPER. Recent studies documenting Env’s inherent stalk flexibility and wide range of ectodomain tilt also suggests looser steric restriction from Env’s soluble domain for an antibody’s bilayer and MPER access than previously thought33,44.

Investigating free antibodies’ membrane-bound conformational preferences, e.g., during maturation or for targeting privileged germlines, may be an informative approach to help illicit bnAbs via vaccine with similar lipid-dependent neutralization mechanism. Thus, we characterized the geometric range of 10E8, 4E10, and PGZL1 ensembles within our simulations via structural clustering and analyzed orientation-dependence of their CDR-phospholipid complexes. Fab surface geometries were defined by immersion depths of each CDR loop and 2 global domain angles relative to membrane plane and thresholded to distinguish 7 substates, distributing well into population fractions from 5-35% (Figure 2–figure supplement 4A-H). Clustering of all bnAb simulation time together expectedly resulted in 4E10 and PGZL1 mapping to shared macroscopic states while the more light-chain directed 10E8 clustered separately (Figure 2–figure supplement 4B-C), reflecting the large differences to CDR-H1 depth and approach angle (Supplementary Table 2). Upon clustering the 4E10 ensemble, interestingly, all substates maintained 86-98% occupancy of the CDR-H1 phospholipid complex whereas 10E8 and PGZL1 ensemble substates showed more conformation-specific variance in their respective phospholipid binding (Figure 2–figure supplement 4E,G,I). For example, two 10E8 substates exhibit low CDR-L1 PC headgroup occupancy (23%, 33%) due to a domain rotation (20-30°) resulting in more CDR-H contacts at the expense of the more typical CDR-L interactions typically observed for 10E8 (Figure 2–figure supplement 4I). These findings highlight the important interplay between bnAbs’ preferred surface orientation and loop composition to form these conformationally sensitive anchoring complexes through positioning of the key phospholipid-binding loops. Loop composition and conformational ensembles are likely honed together during the maturation process.

Antigen influence on membrane bound conformations and lipid binding sites for LN01

We next studied bnAb LN01 to interrogate differences in the antibody surface-bound conformational ensemble and phospholipid interactions in the context of a transmembrane (TM) antigen, utilizing the availability of Fab crystal structures complexed with an MPER-TM fragment (Figure 3)21. MPER-bound LN01 structures bind one phospholipid and one dodecyl-PC detergent in its CDRs, implying possible paratope-epitope-membrane cooperativity in the neutralization mechanism and two putative phospholipid sites anchoring the bnAb to membrane surfaces (Figure 3–figure supplement 1A). The PC headgroup binding site resides in the CDR-H3-L1-L2 inter-chain groove, with the lipid phosphate engaged by L1 Lys31 via salt bridge and the choline moiety interacting in a cation-pi cage with Tyr32 and CDRH3 Tyr100g. When present, MPER contributes additional choline-aromatic stacking (Trp680, Tyr681). The second more solvent-exposed site lies between CDR-H1 and - H2, ordering PS or PC headgroups in crystal structures. We chose to retain the MPER-TM fragment (668-709) as a monomeric continuous helix from the LN01-bound structure21. This conformation has been observed for a similar MPER-TM peptide by NMR45 and within bnAb-bound Env by cryo-EM for TM domains in both crossing (dimeric) or separated (monomeric, “tripod”) arrangements32. We recognize the caveat that the model MPER-TM antigen fragment studied here does not encompass the full structural breadth of MPER display observed or hypothesized for antigen fragments or full-length Env conformational states. However, alternative MPER-TM organizations observed from peptides (e.g. trimeric TM domain helical bundle, kinked MPER in membrane surface-bound or globular trimer)46,47 are less amendable to model due to lack of structural data with bnAbs.

Atomistic simulations of apo and antigen-bound LN01 characterizing paratope-phospholipid complexes with and without epitope and shifted membrane-bound conformation

(A) Representative frame from MD simulation LN01 Fab bound to MPER-TM showing the stable ternary paratope-epitope-membrane complex; bound phospholipids shown.

(B) Frequency of Fab’s characteristic surface-bound geometry by global domain rotation and approach angles in MD simulations for LN01 bound to MPER-TM, plotted by kernel density estimation as contour.

(C) Representative frame from MD simulation of phospholipids complexed with LN01 Fab alone.

(D) Frequency of geometries sampled for apo membrane-bound LN01.

(E) Phospholipid headgroup interaction formed ab initio in LN01+MPER-TM simulations. Aromatic cation-pi cage motifs coordinate choline while the phosphate is coordinating by Lys31 matching the X-ray site binding pose.

(F) The additional distal “Loading” phospholipid site predicted in LN01simulations, with a similar cation-pi cage motif and hydrogen bonds interactions stabilizing the PC headgroup.

(G) Atomic interactions at the X-ray site in apo LN01 simulations.

(H) Interactions at the loading site in apo LN01 simulations.

(I) Lipid headgroup binding in representative simulation of LN01+MPER-TM (n=4 total). Top, X-ray binding site occupancy (green) and phospholipid choline RMSD in a representative trajectory versus experimental position. Bottom, loading site occupancy (cyan) and choline RMSD versus average headgroup bound position.

(J) Lipid headgroup binding for representative apo LN01 trajectory (n=4 total). Site occupancy and RMSD versus predicted binding position for X-ray site (top) and loading site (bottom).

(K) Per-residue interaction profile for MPER-TM-bound LN01 aggregated 1.5 µs from 3 simulations. CDR loops are mapped in solid color blocks below each profile. Fab domain regions making significant contact are labeled, including CDRs and light chain framework region 3 (FR-L3).

(L) Per-residue interaction profile for apo LN01.

Simulations of LN01 were prepared analogously to the other bnAbs with simplified model membranes, replacing 1,2-dioleoyl-sn-glycero-3-phospho-L-serine (DOPS) in place of POPA to provide opportunity to reconstitute LN01’s crystallographic PS binding site. Initial membrane surface geometries for apo and MPER-bound states predicted using PPM21, which accounted for solvation of the MPER-TM fragment, lead to similar embedded configurations having CDR-H3 deeply inserted with surface contacts from framework H3, light chain (CDR-L1, L3), and heavy chain (CDR-H1, L2). Across three 1 μs replicates each, those loops remained inserted and adopt ensembles similar to the starting conformations (Figure 3A-B,K-L). The MPER-antibody interface including key interactions (e.g., Trp100H to Thr676 H-bond) remained stable. The TM domain maintains a continuous helix through MPER without kinking in a stable tilted geometry (minimal rotation about helix axis) that enables the previously observed snorkeling and hydration of TM Arg68648. In both apo and MPER-TM-bound replicates, several POPC binding events were observed within the CDR-H3-L1-L2 groove at atomic accuracy when compared to dodecyl-PC headgroup position in LN01 X-ray structures, with the same sidechain interactions (Figure 3E-J). In contrast to PGZL1, 4E10, and 10E8 simulations, phospholipids in LN01’s CDR-H3-L1-L2 groove site has much lower kinetic stability and occupancy: ∼30% and ∼40% aggregate occupancy for MPER-TM-bound and apo, respectively (Figure 3–figure supplement 1C-D). Occupancy at the second crystallographic CDR-H1-H2 PS/PC site was negligible within any trajectory. Thus, unbiased MD identifies the CDR-H3-L1-L2 groove as the primary phospholipid anchoring site for LN01 in biologically realistic membranes both alone and when bound to TM-embedded MPER, and further emphasizes the importance of lipid in LN01’s maturation and neutralization mechanism.

Two interesting behaviors stood out in LN01 simulations. First, an additional new CDR-phospholipid binding site was observed reproducibly across apo and MPER-TM-bound trajectories at an alternative groove comprised of and CDR-H2, CDR-H3, and FR-L2 we termed the Loading Site (Figure 3E-J). Phospholipids appeared to readily exchange the ∼7 Å distance between this Loading Site and the primary “X-ray” CDR-H3-L1-L2 groove phospholipid binding site (Figure 3–figure supplement 1B). Although, the two sites were often simultaneously occupied. In the Loading Site, lipid headgroup phosphate oxygens were coordinated by hydrogen bonds to LN01 Tyr100g, Tyr49, and Tyr52 – and optionally to gp41 Trp680 if present (Figures 3E-H). Bivalent interaction with the headgroup choline occurs simultaneously via a similar electrostatic interactions and cation-pi cage motif from Tyr49, Tyr52, Thr53, Thr100c, Ser100d, and Tyr100g sidechains. The Loading site was occupied more often than the X-ray site: 78% and 58% simulation time for MPER-TM-bound and apo LN01 replicates respectively (Figures 3I,J; Figure 3-figure supplement 1C-D).

The second notable behavior observed was that antigen-free LN01 sampled a broader range of geometries at the membrane surface, characterized by two angles describing the Fab relative to the membrane (Figure 3A-D). LN01 complexed with MPER-TM was more deeply and extensively embedded than LN01 alone and resulted in a more focused conformation landscape – aligned with the minimal dynamics of the MPER-TM orientation. Protein-lipid interaction profiles reflect consistency in the protein features mediating membrane association between apo and MPER-TM-bound trajectories (Figure 3K-L), with CDR-L and FR-L3 more inserted upon TM antigen engagement upon the minor rearrangement to engage transmembrane antigen. Critically, while apo LN01 favors a different geometry (positive Rotation angle) its membrane surface ensemble heavily samples the major conformation adopted in MPER-TM bound state (negative Rotation angle) (Figure 3B,D). This observation that LN01’s surface-bound geometry is predisposed for transmembrane antigen engagement suggests that maturing bnAbs’ sequences and structures are primed for the membrane interaction key to access MPER and form the neuralization complex.

Coarse-grain ab initio insertion simulations capture biologically relevant membrane-bound conformations

Next, we used extended sampling time afforded by Martini coarse-grained (CG) simulations to more thoroughly explore the process of antibody insertion to lipid bilayers and landscape of feasible bnAb surface-bound conformations49. Fab’s simulated in CG representation with elastic network restraints maintained the protein fold (<2 Å backbone RMSD) and retained key heavy-light inter-domain contacts, informing that this polar bead model rigidly presents the Fab macroscopic surface features critical for mediating bilayer surface interactions, i.e., hydrophobic patches, charge, polarity (Figure 5-figure supplement 1A). Thus, CG simulations might be useful for evaluating bnAbs’ propensity for membrane insertion and approximate preferred conformations, a hypothesis we probed by two distinct simulation approaches.

The first “spontaneous insertion” approach placed a Fab at distinct random initial orientations and distances (0.5 – 2 nm) above a pre-assembled HIV-like anionic lipid bilayer, allowing Fabs free diffusion for 14 µs (Figure 4A). In 18 replicate simulations for each of 4E10, PGZL1, and 10E8, we observed CDR-directed association stable within the lipid bilayer (>1 µs) within 10, 14, and 12 trajectories, respectively (Figure 4B-D, Supplementary Table 3). Simulations captured bilayer surface scanning behavior preceding insertion and numerous dissociation events, pointing to the reversibility and dynamics of the process. As a reference, bovine serum albumin (BSA) was tested similarly. BSA’s documented weak affinity for lipid bilayer association (low millimolar)50 aligns with the drastically reduced membrane contact events (2/18) and lack of sustained insertion observed (Figure 4E). Non-neutralizing anti-gp41 antibody 13h11 having no detectable lipid bilayer interaction by bilayer interferometry7,51 also exhibited negligible insertion and only sparse ‘scanning’ events, starkly contrasting the bnAbs’ behaviors. Thus, these CG simulations readily distinguish MPER bnAbs with high propensity for lipid interaction (low micromolar affinity) from non-specific and low affinity lipid interactions.

Unbiased spontaneous membrane insertion events and semi-biased dissociation events in coarse grain MD simulations

(A) Snapshots of spontaneous insertion event from a Martini model coarse-grain simulation of a 4E10 Fab. The Fab begins in explicit water solvent 0.5-2 nm above a lipid bilayer, freely diffusing and tumbling in bulk solvent, often resulting in a temporary or permanent insertion event (right).

(B,C,D) 18 replicates of coarse grain Fab systems (4E10, PGZL1, 10E8, respectively), initialized with slightly different Fab orientations relative to lipid bilayer. Frames with Fab contacting the membrane are in green and frames with Fab in water (non-associated) are in white for replicate trajectories of 14 µs each.

(E) 18 replicates of coarse grain BSA (top) or 13h11 (bottom) with different starting orientations relative to the lipid bilayer. Membrane contact (green) or diffusion in water (white) shown over 10 µs time.

(F) Snapshots describing a co-assembling membrane pipeline with 4E10 Fab. An Fab is centered in a box with various rotational orientations in space, explicit water, and lipids randomly arranged within a subset of the box (left). By 30ns, the membrane is fully formed (middle). Fab molecules result in a pre-docked membrane bound conformation and sample a permanent insertion event, intermittent membrane association, or dissociation depending on how the Fab contacts with the membrane (right).

(G, H, I) 40 replicates of 5 µs simulations for coarse grain co-assembling systems (4E10, PGZL1, 10E8, respectively), each with slightly different Fab initial orientations relative to lipid bilayer. Membrane contact is classified as above.

A second “co-assembly” CG approach52 aimed to enhance sampling bnAbs insertion and conformations through spontaneous assembly a lipid bilayer around a randomly oriented Fab in a simulation box amongst mixed water, ion, and lipid particles (Figure 4F). Predisposition for Fab-lipid interactions and insertion events were increased, particularly for 10E8 and PGZL1, while time wasted sampling Fab in bulk water was reduced. We combated potential biases of this method for possibly enriching insertion in less favorable conformations and skewing distributions by running more replicates (n=40) of shorter 5 µs simulations. Most trajectories having Fabs interacting with lipid with surface features outside CDRs dissociated quickly or never inserted (Figure 4G-I).

We compared the spectrum of membrane-bound conformations sampled from ab initio CG membrane insertion and with pre-inserted all-atom simulations for each bnAb to characterize correlation. Relative Fab orientation was tracked via two global domain angles: the canonical “angle of approach” (Fab long axis intersecting the membrane normal vector) and an internal “rotation angle” from the vector orthogonal to this long axis (Figure 5A). Fab conformational ensembles and protein-lipid interaction patterns from both CG approaches heavily overlapped with the primary geometry observed for each respective bnAb all-atom trajectories, showing consistency between the two simulation models (Figure 5B, Figure 5-figure supplement 1B-C). CG 4E10 insertions across both methods adopted single conformations which were similar and all-atom-like. PGZL1 and 10E8 showed more promiscuity by sampling alternative ab initio membrane-inserted geometries alongside their respective all-atom-like conformations, with these states appearing consistently in both CG approaches but at different frequencies. The apparent deeper energy well for 4E10’s primary geometry in CG simulations could reflect greater conformational specificity, but could also be a result of the Martini model’s overestimation of hydrophobic interactions53.

Membrane surface-bound bnAb conformations sampled across multiscale simulations

(A) Graphic of angles defined to describe Fab geometries relative to the normal vector at the membrane’s upper leaflet lipid at the phosphate plane in simulation frames (orange arrow). The canonical “approach angle” defines the long axis of the Fab domain (i.e. the central pseudo-symmetry axis) and membrane normal vector (black arrow). A second “rotational angle” is defines the global domain rotation about the Fab pseudo-symmetry axis relative to the membrane normal vector, based on the short axis traversing the light and heavy chains, which is nearly orthogonal to the Fab’s central axis (red arrow).

(B) Frequency plots of rotation and approach angles from frames of membrane-bound Fabs in MD simulations for 4E10 (blue, top row), PGZL1 (red, middle row), and 10E8 (purple, bottom row). Contour plots depicting frequency maxima for angle pairs sampled are by kernel density estimation. Left column, membrane interaction angles sampled from all-atom simulations with Fabs pre-docked using OPM PPM server prediction. Middle column, geometries from coarse-grain membrane co-assembly simulations. Right column, geometries from unbiased spontaneous insertion coarse-grain simulations. Black dots denote the initial Fab-membrane geometries of starting states for replicate trajectories for each antibody initiated in the lipid bilayer.

Overall, these CG simulations facilitate study of ab initio membrane insertion for MPER bnAbs, wherein we find all sustained events are CDR-directed, ruling out contributions from other Fab surface features. For each bnAb, a state we deem to be biologically relevant, based on similarity to geometries observed in all-atom trajectories forming key phospholipid complexes, was either the primary conformation or one of the top 2 sampled. Given that non-CDR directed and alternative CDR-embedded orientations readily dissociate, we conclude that course-grained models can distinguish unfavorable and favorable membrane-bound conformations to an extent that provides utility for characterizing antibody-bilayer interaction mechanisms.

Multi-scale simulations of ab initio formation of bnAb phospholipid complexes

We next integrated these spontaneously inserted CG conformations into all-atom trajectories to track the full ab initio process of bnAb association and determine the competency of CG-derived geometries to subsequently acquire the critical phospholipid anchoring complexes. For each of 4E10, PGZl1, and 10E8, 3 representative medoid Fab poses were extracted by clustering CG simulation frames (clusters A, B, C), and converting to all-atom detail to initiate unbiased MD simulation (Figure 6A-B).

Back-mapping CG membrane-bound geometries to all-atom simulations allows integrative ab initio modeling of the full bnAb insertion process.

(A) Representative frames from membrane-bound 4E10 Fab coarse-grained simulations were back-mapped to all-atom representation to assess the stability and plausibility of those membrane-bound conformations. This coarse-grained-to-all-atom (CG-to-AA) reversion was applied for all medoid frames of interest from each antibody system and used to initiate half-microsecond unbiased all-atom dynamics simulations.

(B) Frequency of membrane interaction angles from coarse grain spontaneous insertion as clustered by geometric substates for 4E10, PGZL1 and 10E8, colored and contoured as in Fig 3B. Corresponding primary all-atom simulations is overlaid (unfilled, black contour frequency density plot).

(C) Conformational geometry sampled upon conversion of CG medoid to an all-atom trajectory. Initial geometry denoted by stars colored matching CG clusters in (B). Frequency and contour plots of conformational angles sampled in stably inserted backmapped all-atom trajectories for 4E10 (left), PGZL1 (middle), and 10E8 (right).

(D) Per-residue interaction profiles for antibody Fab simulations for 4E10 (left), PGZL1 (middle), 10E8 (right) representing each backmapped atomic trajectory, showing CDR-mediated conformations of differing depths and geometries.

(E) Phospholipid headgroup binding and RMSD plots of closest lipid at respective experimentally determined CDR sites for 4E10 (left), PGZL1 (middle), and 10E8 (right), plotted as in Figure 1C or 2B. 4E10 Cluster A is reported in supplement, PGZL1 Cluster A is not reported because no lipid binding was detected, and 10E8 Cluster A is reported in supplement as an artifact.

For 4E10, trajectories initiated from all three geometries drifted back to conformations very similar to those of our initial pre-inserted all-atom trajectories and bound phospholipids at the CDR-H1 site (Figure 6B-E). Interestingly, the trajectory back-mapped from cluster A adopted deeper insertion and higher CDR-H1 phospholipid binding occupancy (>60%) compared to ensembles of clusters B and C trajectories, which were slightly tilted (rotation angle > 0°) and less extensively inserted, indicating the latter geometries are likely less ideal (Figure 6-figure supplement 1B). For PGZL1, all-atom trajectories starting from CG clusters B and C similarly converge to CDR-H-inserted conformations analogous to our previous all-atom MD, and stably bind phospholipid headgroups at CDR-H1 (>50% occupancy) (Figure 6C-E). For 10E8, Cluster C started and finished in membrane-bound orientations analogous to the pre-inserted simulations (Figure 6C); although the Fab adopted a slightly different protein-lipid interaction pattern with much shallower CDR-H3 insertion, the bivalent CDR-L1 phospholipid complex still readily formed with high occupancy (>70%) (Figure 6D-E). Thus, these multi-scale simulations demonstrate that unbiased CG modeling can generate stable membrane-bound conformations ab initio that integrate smoothly into atomistic simulations which are biologically relevant, in that they capture key phospholipid complexes seen in laboratory experiments inferred to be critical for bnAb function.

CG sampling also explored additional alternative membrane-bound geometries, which we assessed for kinetic stability in integrative atomic simulations to discern secondary protein binding modes or intermediate conformations or aiding membrane association (e.g., preliminary surface-scanning) from simple artifacts. The trajectory initiated from PGZL1’s CG cluster A adopts a novel conformation stable on the 500 ns timescale with CDR-H3 inserted, favoring lipid-CDR-L interactions over CDR-H, that was not competent to complex phospholipids at CDR-H1 (Figure 6B-D). All-atom simulation from one alternate conformation of 10E8, CG cluster A, rapidly dissociated from the membrane within 50 ns (Figure 6-figure supplement 1C), showing this integrative MD distinguishes this state as an unstable likely artifact. Interestingly, the all-atom trajectory starting from CG cluster B trended towards the same novel conformation as cluster A, but remained stably inserted and bound phospholipid at the CDR-L1 site at high occupancy (70 %), albeit having shallower CDR-H3 insertion alongside canonical 10E8 CDR-L embedding (Figure 6C-D). These simulations hint at the feasibility to study compatibility of these lipid complex with broader Fab-membrane geometries, and reveal the need for orthogonally evaluating these conformations such as by biased simulations (e.g., potential of mean force, PMF) which we address later in this manuscript.

Additionally, we modelled the orientation of a full-length IgG based on the simulated position of each bnAb Fab to ascertain how sterically feasible a bivalent Fab-membrane interaction could be and possibility of avidity effects in mechanism. The resulting full-length IgG geometries shows that with one membrane-bound Fab, it is very unlikely a second Fabs can simultaneously engage the bilayer due to restriction from the IgG hinge – except for the case an extraordinarily flexible hinge (Figure 6-figure supplement 1D).

Biased pulling simulations distinguish experimentally characterized affinity differences

Finally, we applied biased constant velocity atomistic simulations to compute the force required to dissociate a Fab bound to lipid bilayers – providing a binding strength estimate akin to force spectroscopy (Figure 7A). Doing so tested whether this approach could connect the observed bnAb anchoring phospholipid complexes and membrane-bound conformations with physiological properties such as lipid bilayer affinity and neutralization potency. If informative, this method offers an expedient alternative to canonical free energy calculations such as PMFs that demand tens of µs49, prohibitive for characterizing many antibody variants. Given that Fabs may sample many lipid interactions and surface conformations, we performed ensemble-based measurements: averaging forces from several replicate pulling trajectories with different starting configurations from unbiased MD.

Biased antibody-membrane pulling dissociation simulations approximate Fab-membrane interaction strength for bnAb Fab variants of known lipid binding affinity or neutralization potency.

(A) Schematic of pulling method, a bnAbs Fabs associated with the lipid bilayers is subject to an applied upward dissociation force to the Fab domain center of mass at constant velocity to measure the force required to dissociate the Fab from the bilayer.

(B) The Trp100a-Trp100c motif residues in 4E10 CDR-H3 loop of expected to be lipid-embedded in the bilayer (beige). The double alanine mutant “WAWA” 4E10 variant has experimentally determined significantly reduced affinity to lipid bilayers and lower neutralization potency, due to lack of Trp membrane insertion.

(C) Previously experimentally characterized20 PGZL1 germline-reverted variants, shown as chimera of germline versus matured gene segments, used to approximate antibody properties along its maturation trajectory.

(D) Average force versus distance plots and rupture force (Fmax) calculation for one replicate of pulling wild type 4E10 (WT, blue) and 4E10 WAWA (red) to bias Fab dissociation from the bilayer.

(E) Distribution of rupture forces required for dissociation for different membrane-bound starting conformations for 4E10 (n=9, blue) and 4E10 WAWA (n=11, red), with starting conformations drawn from previous unbiased all-atom simulation ensembles at rest. Outliers as dots

(F) Distributions of rupture forces (n=10) required for membrane dissociation of PGZL1 inferred variants along the maturation pathway, for germline (dark grey, gVgDgJ), intermediate (light grey, gVmDmJ), and mature (maroon, mVmDmJ) PGZL1. A one-tailed two-sample t-test was performed to analyze statistical significance of rupture forces between two antibodies (*=p<0.05; **=p<0.005; ***=p<0.001)

We first assessed suitable pulling velocities through dissociating 4E10 poses from simplified anionic HIV-like lipid bilayers to a fixed 1.5 nm distance over different intervals (10, 50, 100, 200 ns; n=3 unique starting conformations). Interestingly, 2 of 3 starting poses adopt the CDR-H3 rotamer conformation of Trp100b projecting down towards lipid tails (“Trp-Down” state) while the third starting pose has Trp100b in a flipped rotamer pointing towards the lipid-water interface (“Trp-Up” state) (Figure 7-Figure supplement 1A). As expected, rupture forces decrease with increasing simulation time (slower pulling velocity), plateauing between 100 and 200 ns (Figure 7-Figure supplement 1B). Surprisingly, the rupture forces from trajectories starting in the Trp-Down state were nearly identical (< 2% difference), and rupture forces from trajectories starting from the Trp-Up state were consistently lower (27 and 47% greater) for 50 and 100 ns, respectively. This cursory trial exhibited promising conformational sensitivity in distinguishing different protein-lipid interactions in a CDR loop and precision from similar initial Fab conformations.

We next by compared behavior of 4E10 with its well-studied CDR-H3 WAWA variant, the W100aΔA-W100bΔA double mutant (Figure 7B). WAWA has similar Kd to MPER as 4E10, but significantly reduced affinity to empty HIV-like liposomes and neutralization efficacy (>100 fold higher IC50)23,27,30,54. Across replicates from different starting states for 4E10 (n=9) and WAWA (n=11), rupture force calculations showed average forces (± S.E.M.) of 63.8 ± 2.8 kJ/nm/mol and 48.0 ± 2.6 kJ/nm/mol respectively (Figures 7D-E), correctly indicating the 4E10 ensemble has a stronger membrane interaction (32 ± 13 %; p-value < 0.001). Thus, this in silico estimation effectively discerns antibody variants whose high and low lipid bilayer affinities have been previous measured and reinforces the molecular link between WAWA’s reduced membrane binding and neutralization.

Finally, we evaluated experimentally characterized PGZL1 variants that recapitulate possible stages of a germline maturation pathway22 (Figure 7C). Mature PGZL1 has relatively high affinity to the MPER epitope peptide (Kd = 10 nM) and demonstrates great breadth and potency, neutralizing 84% of a 130-strain panel. An “Intermediate” PGZL1 with V-gene germline reversion (gVmDmJ, 100% CDR-H3 identity) showed a modest affinity loss (Kd = 64 nM), while neutralization potency and breadth (12%) were more substantially reduced than expected due to affinity alone – attributed to impaired lipid bilayer interactions22. The fully “Germline” reverted variant (gVgDgM) had no detectable MPER peptide affinity or neutralization. Ensembles for Germline and Intermediate Fab pulling simulations were prepared from unbiased all-atom trajectories, starting from predicted hydrophobicity-optimized inserted poses. The variants adopt insertion geometries like the mature Fab, facilitated by the mostly conserved CDR-H3 (Figure 7-figure supplement 1C), but did not form CDR-H1-phospholipid complexes. Significantly lower pulling forces were required to dissociate Intermediate and Germline PGZL1 from bilayers, 37.5 ± 14.9 kJ/nm/mol and 41.2 ± 11.7 kJ/nm/mol respectively, than for mature PGZL1, 56.1 ± 11.6 kJ/nm/mol (p < 0.003) (Figure 7E). Germline PGZL1 likely reflects kinetically stable but low affinity interactions, representing baseline of rupture forces required to dissociate CDR-inserted antibodies. While the difference in Germline and Intermediate force distributions was not significant (p > 0.4), more higher force events were observed for the Intermediate. These results provide further evidence that bnAb phospholipid binding features are acquired and honed along the maturation pathway, entrenching the connection between positive selection of antibody-lipid interactions and neutralization efficacy.

Discussion

Here, we present a roadmap for applying multi-scale molecular simulations to characterize the biophysical underpinning of lipid membrane interactions within the mechanism of HIV MPER bnAbs. The approaches demonstrated and principles discerned improve understanding of maturation pathways for antibodies targeting membrane-proximal epitopes and should enrich data-driven design of HIV immunogens. Our detailed simulations supplement in vitro binding and crystallographic evidence in establishing that bnAbs develop highly specific phospholipid interactions that facilitate access to the MPER epitope and can participate in epitope-paratope interface in context of full lipid bilayers. These results add validation to which out of the many possible ions and short-chain lipid moieties ordered within those crystal structures may be inferred as putative bilayer phospholipids, a critical consideration in antibody and vaccine design. The additional simulations outlined with different lipid ratios and varied antigens as well as more rigorous PMF studies, could be applied to better inform affinity thresholds and energetic relationships paramount to the lipid complexes and conformational preferences reported here. Furthermore, these modeling approaches should have utility in finding membrane-targeting elements in antibodies towards identifying causes of autoimmune stress. Likewise, the analysis tools in this work might be practical for aiding proactive in antibody engineering targeting membrane proteins, for example grafting lipid phosphate binding structural motifs or amino acid sequences with desirable membrane associating properties into distinct antibody loops regions.

Further, these simulations can serve future examination of molecular details concerning the genetic origins and developmental pathways for incipient lipid-binding antibodies. Through natural or vaccine-induced immunity, coaxing the immune system to develop both lipid and antigen affinity (possibly even cooperativity) during maturation is a difficult task, and hindered by downregulation of membrane binding precursors1,3,6,11. Likewise, MPER bnAbs are difficult to induce and, until recent clinical trials, have only been isolated from patients with chronic HIV-infection and sustained immunosuppression13,13,14,19,20,38. Modern vaccine strategies often elicit certain subsets of precursor B cells, often targeting specific germline genes, in attempt to guide antibodies’ mature of specific molecular features intended for the targeted epitopes55,56. The question remains whether particular germline genes and lineages are privileged for successful maturation as membrane-targeting (or MPER-targeting) antibodies. This notion is supported by 4E10 and PGZL1 sharing a germline gene (VH1-69)22 and nearly identical membrane interactions. Certain germlines genes may be predisposed with inherent basal membrane affinity or be favored precursor scaffolds for compatibility with evolution of lipid-binding properties. Specific CDR lipid-binding motifs are predicted to be encoded early in maturation processes for PGZL122, namely the CDR-H1 loop, and complemented by mutations at membrane-contacting FR-H3 and CDR-H3 loop residues incorporated later in maturation (Figure 7-figure supplement 1C). Across MPER bnAbs, membrane-binding molecular features are likely acquired at strategic timepoints throughout development to balance poly-specificity and evade autoimmune checkpoints57. Experimental databases of germlines genes usage and of B-cell repertoires sequenced during immunization courses would be ideally paired with the simulations described here to investigate in vivo filtering of lineages with lipid-binding variants and assess possible rules of the autoimmune system58.

Beyond gp41-targeting antibodies and lipid antibodies in autoinflammatory diseases59,60 or microbial infections61,62, we suspect a broader scope of positive outcomes from antibody tolerance and interaction with lipid bilayers may occur in Nature thus is currently underestimated and underutilized. Although host cross-specificity of 4E10 is well documented, the vastly reduced poly-reactivity of PGZL1 and 10E8 inspires optimism that maturation of phospholipid interactions and membrane tolerance may be incorporated into strategies for vaccine design and therapeutic antibodies for other membrane proteins with conserved buried epitopes36. Currently, the breadth of analogous antibodies targeting membrane proteins via extensive or cooperative lipid interactions is poorly explored. The simulation procedures and principles described here are well positioned to investigate this outstanding question and could help define broader chemical rules for design of membrane-interfacing antibodies.